---
title: "Untitled"
output: html_document
date: "2025-07-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Genomic values
```{r}
gv <- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp_out.gv")
gv <- gv |> dplyr::rename(GV = V1)
samples <- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp.fam")
samples <- samples[, c(1, 6)]
samples <- samples |> dplyr::rename(ID = V1, Survival = V6)

gv2<- cbind(samples, gv)
adults <- gv2[175:217, ]
adults<- adults |> dplyr::select(ID, GV)

larvae <- gv2[1:174, ]
```

Larvae breeding values
```{r}
larvae$ID <- sub("-Cn.*", "", larvae$ID)
larvae$ID <- sub(".*-", "", larvae$ID)
larvae$reefcross <- gsub("[0-9]", "", larvae$ID)

larvae |> ggplot() +
  geom_boxplot(aes(x = reefcross, y = GV)) +
  geom_point(aes(x = reefcross, y = GV))
```

Adult breding values
```{r}
adults$ID <- sub(".*A\\.ten-", "", adults$ID)
adults$Reef <- sub("-.*", "", adults$ID)

adults |> ggplot() +
  geom_boxplot(aes(x = Reef, y = GV)) +
  geom_point(aes(x = Reef, y = GV))
```

Heritability estimate
```{r}
hyp <- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp_out.hyp", header=TRUE)
gen_var <- hyp$Va  # Va: additive genetic variance
res_var <- hyp$Ve  # Ve: residual variance
h2 <- gen_var / (gen_var + res_var)

summary(h2)
hist(h2, main="Posterior distribution of heritability", xlab="h²")

#It tells you what proportion of phenotypic variance is explained by all the observed SNPs in your dataset.

model <- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp_out_model.txt", header=FALSE)

h2 <- 1.520914e+02 / (1.520914e+02 + 1.219891e+02)
```

Proportion of phenotypic variance explained by SNPs

This is the same as the heritability estimate in BayesR, because it only uses SNP effects (i.e., narrow-sense h², not total broad-sense h²).
It tells you what proportion of phenotypic variance is explained by all the observed SNPs in your dataset.

.Param SNP effect size
```{r}
param<- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp_out_param.txt", header = TRUE)
SNPs<- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp.bim", header = FALSE)
param2<- cbind(SNPs, param)

hist(param2$alpha, breaks = 10)
threshold <- quantile(abs(param2$alpha), 0.95)
param2$nonzero_pip <- param2$PIP2 + param2$PIP3 + param2$PIP4
important_snps <- param2[param2$nonzero_pip > 0.2 & abs(param2$alpha) > threshold, ]
```

```{r}
hyp <- read.table("/Users/alexmacadam/Desktop/bayesR/aten_larvae_T0_adult_imp_out.hyp", header=TRUE)
h2 <- hyp$Va / (hyp$Va + hyp$Ve)

# Summarize or plot it:
summary(h2)
hist(h2, main="Posterior distribution of heritability", xlab="h²")

#uncertainty in variance
quantile(h2, probs = c(0.025, 0.5, 0.975))  # 95% credible interval

```

```{r}
# Average number of SNPs in each class
colMeans(hyp[, grep("^Nk", names(hyp))])

# Average variance explained by each class
colMeans(hyp[, grep("^Vk", names(hyp))])

library(reshape2)
nk_data <- hyp[, grep("^Nk", names(hyp))]
nk_melt <- reshape2::melt(nk_data, variable.name="Mixture", value.name="Count")
boxplot(Count ~ Mixture, data=nk_melt, main="SNP Counts in Mixture Components")

plot(hyp$Nk4, type = "l", main = "Traceplot: Number of SNPs in High-Effect Class (Nk4)")
```

