---
title: "02_A.ten_larvae_diversity"
output: html_document
date: "2024-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

alpha= within popualtion, eg heterozygosity, allelic richness, Nucleotide diversity
beta= between populations, eg, Fst, AMOVA, Jaccard Index, Bray curtis

#calculate alpha and beta diversity for q=0, q=1 and q=2
#within population diversity, i.e. alpha values:
#q=0 represents allelic richness index
#q=1 Shannon information index
#q=2 heterozygosity index
#indices are represented by H values. These H values can then be converted to a common scale of effective numbers, D.
#For between population comparisons of diversity, i.e. beta values:
#q=0 represents number of allelic types not shared between pops
#q=1 measures the chance of a type of individuals providing #info on the sampling locality
#q=2 measured the chance of two random individuals from different localities having different allelic types

##############
load packages
```{r}
set.seed(123)
library(dartRverse)
library(tidyverse)
library(adegenet)
library(poppr)
library(hierfstat)
library(reshape2)
library(RColorBrewer)
library(scales)
library(LEA)
library(wesanderson)
library(tibble)
library(purrr)
library(ggthemes)
library(vegan)
```

Load personal theme
```{r}
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

#pallets
pd <- position_dodge(width=0.2)
pal <- wes_palette("Zissou1", 5, type = "discrete")
scalesize <- function(x) sprintf("%.1f", x)
pal2 <- c("#386cb0","#7fc97f","#fdb462")
```

PERCENT SURVIVAL function
```{r}
# Function for summarising survival standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
  
  #Use the function like this:
  #summarySE(data, measurevar="Value", groupvars=c("Metric1","Metric2"), na.rm=TRUE)
}
```

load meta data
```{r}
#load metadata
meta<- read.csv("/Users/alexmacadam/Documents/GitHub/AGF-DArT/data/aten_metadata.csv") |>
  filter(lifestage == "larvae")

#remove models and summarise
sum<- meta |> filter(Temp == "T0", !family %in% c("Mod1","Mod2", "Mod3", "Mod4","Mod5","Mod6")) |>
  group_by(pop, indcross) |>
  dplyr::summarise(n = n())
```

Load data
```{r}
#dart data
ten_larvae <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/01_DArT_files/01_DArT_data_filtered/ten_larvae_filter.RDS")
ten_larvae <- gl.drop.pop(ten_larvae, pop.list=c('ParxDav_T0', 'ParxDav_Amb', 'ParxDav_Hot'),mono.rm=FALSE,recalc=TRUE)
```

Remove model samples
```{r}
ten_larvae<- gl.drop.ind(ten_larvae, ind.list = c("Mod1-JewBotxDav6Bul-Cn1-T0-rep1","Mod2-Par6BulxPar6Bul-Cn3-T0-rep3","Mod3-Par6BulxDav6Bul-Cn5-T0-rep5","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep1","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep3","Mod1-JewBotxDav6Bul-Cn1-T0-rep2","Mod2-Par6BulxPar6Bul-Cn3-T0-rep4","Mod3-Par6BulxDav6Bul-Cn5-T0-rep6","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep2","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep4","Mod1-JewBotxDav6Bul-Cn1-T0-rep3","Mod2-Par6BulxPar6Bul-Cn3-T0-rep5","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep1","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep3","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep5","Mod1-JewBotxDav6Bul-Cn1-T0-rep4","Mod2-Par6BulxPar6Bul-Cn3-T0-rep6","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep2","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep4","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep6","Mod1-JewBotxDav6Bul-Cn1-T0-rep5","Mod3-Par6BulxDav6Bul-Cn5-T0-rep1","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep3","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep5","Mod1-JewBotxDav6Bul-Cn1-T0-rep6","Mod3-Par6BulxDav6Bul-Cn5-T0-rep2","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep4","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep6","Mod2-Par6BulxPar6Bul-Cn3-T0-rep1","Mod3-Par6BulxDav6Bul-Cn5-T0-rep3","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep5","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep1","Mod2-Par6BulxPar6Bul-Cn3-T0-rep2","Mod3-Par6BulxDav6Bul-Cn5-T0-rep4","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep6","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep2","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep5","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep3","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep5","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep1","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep5","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep6","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep4","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep6","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep2","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep6","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep5","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep3","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep1","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep6","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep4","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep2","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep1","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep5","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep1","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep3","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep1","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep2","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep6","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep2","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep4","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep2","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep1","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep3","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep3","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep5","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep3","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep2","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep4","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep4","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep6","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep4","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep3","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep1","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep3","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep2","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep4","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep2","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep4","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep3","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep5","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep3","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep5","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep4","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep2","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep6","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep4","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep6","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep5","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep3","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep5","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep6","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep6","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep1","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep2","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep2"), mono.rm = FALSE, recalc=TRUE)
```

Create Pre, Hot and Ambient gl objects
```{r}
ten_larvae_Amb<- gl.keep.pop(ten_larvae, pop.list = c('Amb'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_Hot<- gl.keep.pop(ten_larvae, pop.list = c('Hot'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_T0<- gl.keep.pop(ten_larvae, pop.list = c('T0'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
```

#############
By reef
50 iterations of diversity metrics
Pre
```{r}
# Check how many samples per population
table(pop(ten_larvae_T0))

# Define the minimum sample size and number of draws
min_sample_size_T0 <- 12
num_draws_T0 <- 50

# Initialize a list to store results for each draw
subsample_results_T0 <- vector("list", num_draws_T0)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_T0) {
  subsample_results_T0[[i]] <- gl.subsample.ind(
    x = ten_larvae_T0,          # The genlight object
    n = min_sample_size_T0,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_T0[[1]]))

# Calculate diversity metrics for each draw
diversity_results_T0 <- lapply(subsample_results_T0, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_T0 <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_T0 <- diversity_results_T0[[i]]
  
  # Create a dataframe with the relevant columns
  df_T0 <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_beta = iteration_data_T0$zero_H_beta,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_T0[[i]] <- df_T0
}

# Combine all the dataframes into one
diversity_combined_T0 <- do.call(rbind, diversity_summary_list_T0) |>
  rownames_to_column(var = "population") |>
  plyr::mutate(Reefcross = substr(population, 1, 7))|>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_T0<- rbind(long_ar_T0, long_SI_T0, long_He_T0)

# Calculate SE
long_diversity_T0<- long_diversity_T0 |> mutate(se = sd/sqrt(48))

diversity_sum_T0 <- summarySE(long_diversity_T0, measurevar="Value",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::select(Metric, Reefcross, mean)

diversity_se_T0 <- summarySE(long_diversity_T0, measurevar="se",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_pre <- cbind(diversity_sum_T0, diversity_se_T0) |>
  mutate(Temp = "Pre")
```

```{r}
#save one dataframe for now
beta<- diversity_summary_list_T0[[1]]
```

```{r}
#pivot longer
beta2<- beta |>
  rownames_to_column(var = "cross1") |>
  dplyr::select(-Iteration) |>
  pivot_longer(cols = c(2:10), names_to = "cross2", values_to = "gen_dif") |>
  na.omit()
# Remove "zero_H_beta." from every cell of the 'cross2' column in the beta2 dataframe
beta2$cross2 <- gsub("zero_H_beta.", "", beta2$cross2)
# Add comparison column to 'beta2' with columns 'column1' and 'column2'
beta2$comparison <- paste(beta2$cross1, beta2$cross2, sep = "_vs_")
```

```{r}
#list of comparisons to keep
keep<- c('DavxDav_T0_vs_JewxDav_T0','PalxPal_T0_vs_JewxPal_T0','DavxDav_T0_vs_JewxJew_T0','PalxPal_T0_vs_JewxJew_T0','DavxDav_T0_vs_ParxPar_T0','DavxDav_T0_vs_SwixDav_T0','DavxDav_T0_vs_SwixSwi_T0')

beta3<- beta2 |>
  dplyr::filter(comparison %in% keep)
beta3<- beta3 |>
  mutate(comparison = factor(comparison, levels = c('DavxDav_T0_vs_JewxDav_T0','PalxPal_T0_vs_JewxPal_T0','DavxDav_T0_vs_JewxJew_T0','PalxPal_T0_vs_JewxJew_T0','DavxDav_T0_vs_ParxPar_T0','DavxDav_T0_vs_SwixDav_T0','DavxDav_T0_vs_SwixSwi_T0')))
```

```{r}
#juvenile response
response<- c(1.697448311,2.781274217,6.015635249,7.001579956,4.626996881,1.114827565,-2.4376347)
responses<- as.data.frame(cbind(keep, response)) |>
  dplyr::rename(comparison = keep)

beta4<- beta3 |>
  left_join(responses, by = "comparison")

beta4$response <- as.numeric(beta4$response)

beta4<- beta4 |>
  mutate(ratio = response/gen_dif) |>
  mutate(log_ratio = log(response/gen_dif))

beta4$comparison <- gsub("_T0", "", beta4$comparison)
```

```{r}
cor1<- beta4 |>
  ggplot(aes(x = gen_dif, y = response)) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  geom_point(aes(colour = comparison), size = 2) +
  #geom_point(aes(x = 0.2807432, y = -2.437635)) +
  labs(x = "Genetic difference", y = "Selection Response", colour = "Reef Cross Comparison") +
  theme_classic() +
  stat_cor(label.x = 0.25, label.y = 6)

cor2<- beta4 |> filter(comparison != "DavxDav_vs_SwixSwi") |>
  ggplot(aes(x = gen_dif, y = response)) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  geom_point(aes(colour = comparison), size = 2) +
  labs(x = "Genetic difference", y = "Selection Response", colour = "Reef Cross Comparison") +
  theme_classic() +
  stat_cor(label.x = 0.25, label.y = 6) 
  #stat_regline_equation(label.x = 0.25, label.y = 5)

library(patchwork)
cor<- cor1 + cor2
```
