---
title: "06_A.ten_larvae_survival_correlation"
output: html_document
date: "2024-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and tools
```{r}
library(tidyverse)
library(ggpubr)
library(dartR)
library(wesanderson)
library(janitor)
library(glmnet)

set.seed(123)

#colour palette
pal <- wes_palette("Zissou1", 5, type = "discrete")

#theme for plots
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "right",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               #strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold"),
               strip.background = element_rect(fill = "white", colour = "black", size = rel(2))
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#3b9ab2","#f21a00","#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33","grey")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#3b9ab2","#f21a00","#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33","grey")), ...)

}

# Function for summarising survival standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

Load data
```{r}
#survival data
survival<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/aten_larvae_survival_corr.csv")
#dart data
ten_larvae <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_larvae_filter.RDS")
#keep T0
ten_larvae_T0<- gl.keep.pop(ten_larvae, pop.list = c('T0'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
#drop models
ten_larvae_T0<- gl.drop.pop(ten_larvae_T0, pop.list = c('Mod1','Mod2','Mod3','Mod4','Mod5','Mod6'), as.pop = 'family', mono.rm = FALSE, recalc=TRUE)
```

Create objects for each family
```{r}
Dav1xDav8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav1xDav8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Dav6xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav6xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Dav8xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav8xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew2xJew8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew2xJew8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew2xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew2xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew3xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew3xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew6xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew6xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav10 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav10"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav13 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav13"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xJew2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xJew2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xJew3 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xJew3"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal2xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal2xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal2xPal5 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal2xPal5"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal4xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal4xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal4xPal2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal4xPal2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal5xPal4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal5xPal4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par2xPar5 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par2xPar5"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par4xPar2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par4xPar2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par5xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par5xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par5xPar4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par5xPar4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi4xDav8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi4xDav8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi4xSwi7 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi4xSwi7"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi6xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi6xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi6xSwi4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi6xSwi4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi7xDav10 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi7xDav10"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi7xSwi6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi7xSwi6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
```

Create dataframe of allele frequency for each family
```{r}
Dav1xDav8_alf <- gl.alf(Dav1xDav8) |> rownames_to_column() |> dplyr::rename(Dav1xDav8 = alf1) |> dplyr::select(Dav1xDav8)
Dav6xDav1_alf <- gl.alf(Dav6xDav1) |> rownames_to_column() |> dplyr::rename(Dav6xDav1 = alf1) |> dplyr::select(Dav6xDav1)
Dav8xDav6_alf <- gl.alf(Dav8xDav6) |> rownames_to_column() |> dplyr::rename(Dav8xDav6 = alf1) |> dplyr::select(Dav8xDav6)
Jew2xJew8_alf <- gl.alf(Jew2xJew8) |> rownames_to_column() |> dplyr::rename(Jew2xJew8 = alf1) |> dplyr::select(Jew2xJew8)
Jew2xPal1_alf <- gl.alf(Jew2xPal1) |> rownames_to_column() |> dplyr::rename(Jew2xPal1 = alf1) |> dplyr::select(Jew2xPal1)
Jew3xPal1_alf <- gl.alf(Jew3xPal1) |> rownames_to_column() |> dplyr::rename(Jew3xPal1 = alf1) |> dplyr::select(Jew3xPal1)
Jew6xPal1_alf <- gl.alf(Jew6xPal1) |> rownames_to_column() |> dplyr::rename(Jew6xPal1 = alf1) |> dplyr::select(Jew6xPal1)
Jew8xDav1_alf <- gl.alf(Jew8xDav1) |> rownames_to_column() |> dplyr::rename(Jew8xDav1 = alf1) |> dplyr::select(Jew8xDav1)
Jew8xDav10_alf <- gl.alf(Jew8xDav10) |> rownames_to_column() |> dplyr::rename(Jew8xDav10 = alf1) |> dplyr::select(Jew8xDav10)
Jew8xDav13_alf <- gl.alf(Jew8xDav13) |> rownames_to_column() |> dplyr::rename(Jew8xDav13 = alf1) |> dplyr::select(Jew8xDav13)
Jew8xDav6_alf <- gl.alf(Jew8xDav6) |> rownames_to_column() |> dplyr::rename(Jew8xDav6 = alf1) |> dplyr::select(Jew8xDav6)
Jew8xJew2_alf <- gl.alf(Jew8xJew2) |> rownames_to_column() |> dplyr::rename(Jew8xJew2 = alf1) |> dplyr::select(Jew8xJew2)
Jew8xJew3_alf <- gl.alf(Jew8xJew3) |> rownames_to_column() |> dplyr::rename(Jew8xJew3 = alf1) |> dplyr::select(Jew8xJew3)
Jew8xPal1_alf <- gl.alf(Jew8xPal1) |> rownames_to_column() |> dplyr::rename(Jew8xPal1 = alf1) |> dplyr::select(Jew8xPal1)
Pal2xDav1_alf <- gl.alf(Pal2xDav1) |> rownames_to_column() |> dplyr::rename(Pal2xDav1 = alf1) |> dplyr::select(Pal2xDav1)
Pal2xPal5_alf <- gl.alf(Pal2xPal5) |> rownames_to_column() |> dplyr::rename(Pal2xPal5 = alf1) |> dplyr::select(Pal2xPal5)
Pal4xDav1_alf <- gl.alf(Pal4xDav1) |> rownames_to_column() |> dplyr::rename(Pal4xDav1 = alf1) |> dplyr::select(Pal4xDav1)
Pal4xPal2_alf <- gl.alf(Pal4xPal2) |> rownames_to_column() |> dplyr::rename(Pal4xPal2 = alf1) |> dplyr::select(Pal4xPal2)
Pal5xPal4_alf <- gl.alf(Pal5xPal4) |> rownames_to_column() |> dplyr::rename(Pal5xPal4 = alf1) |> dplyr::select(Pal5xPal4)
Par2xPar5_alf <- gl.alf(Par2xPar5) |> rownames_to_column() |> dplyr::rename(Par2xPar5 = alf1) |> dplyr::select(Par2xPar5)
Par4xPar2_alf <- gl.alf(Par4xPar2) |> rownames_to_column() |> dplyr::rename(Par4xPar2 = alf1) |> dplyr::select(Par4xPar2)
Par5xDav6_alf <- gl.alf(Par5xDav6) |> rownames_to_column() |> dplyr::rename(Par5xDav6 = alf1) |> dplyr::select(Par5xDav6)
Par5xPar4_alf <- gl.alf(Par5xPar4) |> rownames_to_column() |> dplyr::rename(Par5xPar4 = alf1) |> dplyr::select(Par5xPar4)
Swi4xDav8_alf <- gl.alf(Swi4xDav8) |> rownames_to_column() |> dplyr::rename(Swi4xDav8 = alf1) |> dplyr::select(Swi4xDav8)
Swi4xSwi7_alf <- gl.alf(Swi4xSwi7) |> rownames_to_column() |> dplyr::rename(Swi4xSwi7 = alf1) |> dplyr::select(Swi4xSwi7)
Swi6xDav1_alf <- gl.alf(Swi6xDav1) |> rownames_to_column() |> dplyr::rename(Swi6xDav1 = alf1) |> dplyr::select(Swi6xDav1)
Swi6xSwi4_alf <- gl.alf(Swi6xSwi4) |> rownames_to_column() |> dplyr::rename(Swi6xSwi4 = alf1) |> dplyr::select(Swi6xSwi4)
Swi7xDav10_alf <- gl.alf(Swi7xDav10) |> rownames_to_column() |> dplyr::rename(Swi7xDav10 = alf1) |> dplyr::select(Swi7xDav10)
Swi7xSwi6_alf <- gl.alf(Swi7xSwi6) |> rownames_to_column() |> dplyr::rename(Swi7xSwi6 = alf1) |> dplyr::select(Swi7xSwi6)
```

extract SNP names
```{r}
SNPs <- gl.alf(Dav1xDav8) |> rownames_to_column() |> dplyr::rename(SNPs = rowname) |> dplyr::select(SNPs)
```

Bind all allele frequencies together
```{r}
alf<- cbind(
SNPs,
Dav1xDav8_alf$Dav1xDav8, 
Dav6xDav1_alf$Dav6xDav1, 
Dav8xDav6_alf$Dav8xDav6, 
Jew2xJew8_alf$Jew2xJew8, 
Jew2xPal1_alf$Jew2xPal1, 
Jew3xPal1_alf$Jew3xPal1, 
Jew6xPal1_alf$Jew6xPal1, 
Jew8xDav1_alf$Jew8xDav1, 
Jew8xDav10_alf$Jew8xDav10, 
Jew8xDav13_alf$Jew8xDav13, 
Jew8xDav6_alf$Jew8xDav6, 
Jew8xJew2_alf$Jew8xJew2, 
Jew8xJew3_alf$Jew8xJew3, 
Jew8xPal1_alf$Jew8xPal1, 
Pal2xDav1_alf$Pal2xDav1, 
Pal2xPal5_alf$Pal2xPal5, 
Pal4xDav1_alf$Pal4xDav1, 
Pal4xPal2_alf$Pal4xPal2, 
Pal5xPal4_alf$Pal5xPal4, 
Par2xPar5_alf$Par2xPar5, 
Par4xPar2_alf$Par4xPar2, 
Par5xDav6_alf$Par5xDav6, 
Par5xPar4_alf$Par5xPar4, 
Swi4xDav8_alf$Swi4xDav8, 
Swi4xSwi7_alf$Swi4xSwi7, 
Swi6xDav1_alf$Swi6xDav1, 
Swi6xSwi4_alf$Swi6xSwi4, 
Swi7xDav10_alf$Swi7xDav10, 
Swi7xSwi6_alf$Swi7xSwi6
)

rm(Dav1xDav8_alf,Dav1xDav8, 
Dav6xDav1_alf,Dav6xDav1, 
Dav8xDav6_alf,Dav8xDav6, 
Jew2xJew8_alf,Jew2xJew8, 
Jew2xPal1_alf,Jew2xPal1, 
Jew3xPal1_alf,Jew3xPal1, 
Jew6xPal1_alf,Jew6xPal1, 
Jew8xDav1_alf,Jew8xDav1, 
Jew8xDav10_alf,Jew8xDav10, 
Jew8xDav13_alf,Jew8xDav13, 
Jew8xDav6_alf,Jew8xDav6, 
Jew8xJew2_alf,Jew8xJew2, 
Jew8xJew3_alf,Jew8xJew3, 
Jew8xPal1_alf,Jew8xPal1, 
Pal2xDav1_alf,Pal2xDav1, 
Pal2xPal5_alf,Pal2xPal5, 
Pal4xDav1_alf,Pal4xDav1, 
Pal4xPal2_alf,Pal4xPal2, 
Pal5xPal4_alf,Pal5xPal4, 
Par2xPar5_alf,Par2xPar5, 
Par4xPar2_alf,Par4xPar2, 
Par5xDav6_alf,Par5xDav6, 
Par5xPar4_alf,Par5xPar4, 
Swi4xDav8_alf,Swi4xDav8, 
Swi4xSwi7_alf,Swi4xSwi7, 
Swi6xDav1_alf,Swi6xDav1, 
Swi6xSwi4_alf,Swi6xSwi4, 
Swi7xDav10_alf,Swi7xDav10, 
Swi7xSwi6_alf,Swi7xSwi6
)
```

Transpose and create rownames
```{r}
alf<- alf |> 
  t() |>
  row_to_names(row_number = 1)
```

Change class of matrix
```{r}
class(alf) <- "numeric"
alf2<- as.data.frame(alf)
```

Correlation between SNP alf and Larval survival
```{r}
#run correlation analysis
cor1<-psych::corr.test(alf2[c(1:13057)], survival$Survival, method="spearman", adjust = "BH")
ci<- cor1$ci2
padj<- cor1$p.adj
```

Make datframes of highly significant SNPs
```{r}
#filter highly predictive
highly_pred_SNPs<- ci |> dplyr::filter(pa < 0.05)

#extract SNP names (ci for some reason butchers the names)
pearson_snps_0.5<- highly_pred_SNPs |> rownames_to_column(var = "SNP")
pearson_snps_0.5<- pearson_snps_0.5$SNP
pearson_snps_0.5<- gsub("-NA", "", pearson_snps_0.5)
SNP_list<- SNPs$SNPs |> as.vector()
pattern <- paste(pearson_snps_0.5, collapse = "|")
pearson_snps_0.5<- SNP_list[grep(pattern, SNP_list)]
```

Plot correlations
```{r}
#plot positive correlations
SNP_15476258<- alf2 |> dplyr::select(`15476258-41-A/G`)
SNP_15476258<- cbind(SNP_15476258, survival)
plot15476258<- ggplot(SNP_15476258, aes(x = `15476258-41-A/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='glm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15476258

SNP_15477968<- alf2 |> dplyr::select(`15477968-19-C/A`)
SNP_15477968<- cbind(SNP_15477968, survival)
plot15477968<-ggplot(SNP_15477968, aes(x = `15477968-19-C/A`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15477968

SNP_73055070<- alf2 |> dplyr::select(`73055070-13-A/G`)
SNP_73055070<- cbind(SNP_73055070, survival)
plot73055070<-ggplot(SNP_73055070, aes(x = `73055070-13-A/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot73055070

SNP_15457915<- alf2 |> dplyr::select(`15457915-49-G/C`)
SNP_15457915<- cbind(SNP_15457915, survival)
plot15457915<-ggplot(SNP_15457915, aes(x = `15457915-49-G/C`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15457915

#plot negative correlations
SNP_9498501<- alf2 |> dplyr::select(`9498501-47-C/T`)
SNP_9498501<- cbind(SNP_9498501, survival)
plot9498501<-ggplot(SNP_9498501, aes(x = `9498501-47-C/T`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot9498501

SNP_11466799<- alf2 |> dplyr::select(`11466799-26-T/G`)
SNP_11466799<- cbind(SNP_11466799, survival)
plot11466799<-ggplot(SNP_11466799, aes(x = `11466799-26-T/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot11466799

SNP_11461901<- alf2 |> dplyr::select(`11461901-9-T/C`)
SNP_11461901<- cbind(SNP_11461901, survival)
plot11461901<- ggplot(SNP_11461901, aes(x = `11461901-9-T/C`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot11461901

#combine plots
library(patchwork)
plot15476258 + plot15477968 + plot73055070 + plot9498501 + plot11466799 + plot11461901 + plot_layout(ncol = 3, guides = 'collect') + plot_annotation(tag_levels = 'A')
```

Find and plot significant SNPs to see differences in AF between region crosses
```{r}
#filter out the significant SNPs
sig_als<- padj |> as.data.frame() |> dplyr::filter(V1 < 0.05) |> rownames_to_column(var = "SNPs")
sig_als <- sig_als$SNPs |> as.vector()
sig_als2<- alf2 |> dplyr::select(all_of(sig_als))

#bind survival and allele frequency together
sig_als3<- cbind(survival, sig_als2)
rownames(sig_als3) <- NULL

#Create dataframe of sig NSP AF per region
summary_AF <- sig_als3 |>
  group_by(Region) |>
  dplyr::summarise(across(contains("/"), mean, na.rm = TRUE)) |>
  pivot_longer(cols = contains("/"), 
               names_to = "SNP", 
               values_to = "allele_frequency")

# Plot violin plot of allele frequencies by region
vio_af_plot<- ggplot(summary_AF, aes(x = Region, y = allele_frequency, colour = Region)) +
  geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") +
  scale_color_manual(values=pal[c(1,3,5)]) +
  scale_fill_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.direction = "vertical")
vio_af_plot

hist_plot<- ggplot(summary_AF) +
  geom_histogram(aes(x =allele_frequency, fill = Region)) +
  facet_wrap(~Region) +
  labs(x = "Allele Frequency", y = "Count") +
  scale_color_manual(values=pal[c(1,3,5)]) +
  scale_fill_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.direction = "vertical")
hist_plot
```

Line plot of SNPs
```{r}
#how many are fixed in central corals?
fixed<- summary_AF |>
  filter(Region == "Central") |>
  filter(allele_frequency < 0.01 | allele_frequency > 0.99) |>
  pull(SNP)

# Calculate the standard deviation of allele frequencies of the Central and North larvae
variation_summary <- summary_AF %>% dplyr::filter(Region != "Hybrid") |>
  group_by(SNP) %>%
  dplyr::summarise(var = sd(allele_frequency)) %>%
  dplyr::arrange(desc(var))

# Filter for the top SNPs with highest variation by region
top_SNPs <- variation_summary %>%
  filter(var > quantile(var, 0.75)) %>%  # Keep SNPs with variation in the top 25%
  pull(SNP)

# Calculate the standard deviation of allele frequencies by SNP across regions
#top_SNPs <- summary_AF |>
#  dplyr::filter(Region == "Central") |>
#  group_by(SNP) |>
#  dplyr::filter(allele_frequency > 0.95 | allele_frequency < 0.05) |>
#  pull(SNP)

# Filter the original data for top SNPs
highlighted_data <- summary_AF %>% 
  filter(SNP %in% fixed)

# Plot with highlighted SNPs
lin_af_plot<-ggplot(highlighted_data, aes(x = Region, y = allele_frequency, colour = Region, group = SNP)) + 
  #geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_line(aes(group = SNP), colour = "gray", size = 0.8, alpha = 0.5) +  # Lines connecting points for same SNP
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1.5) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") + 
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.direction = "vertical")
lin_af_plot
```

```{r}
# Plot violin plot of allele frequencies by region
vio_af_plot<- ggplot(summary_AF, aes(x = Region, y = allele_frequency, colour = Region)) +
  geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") +
  scale_color_manual(values=pal[c(1,3,5)]) +
  scale_fill_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
vio_af_plot

# Plot with highlighted SNPs
lin_af_plot<-ggplot(highlighted_data, aes(x = Region, y = allele_frequency, colour = Region, group = SNP)) + 
  #geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_line(aes(group = SNP), colour = "gray", size = 0.8, alpha = 0.5) +  # Lines connecting points for same SNP
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1.5) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") + 
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lin_af_plot

#combined plot
af_plot<- ggplot(summary_AF, aes(x = Region, y = allele_frequency, colour = Region)) +
  
  # Lines for highlighted SNPs
  geom_line(
    data = highlighted_data,
    aes(x = Region, y = allele_frequency, group = SNP),
    colour = "gray", size = 0.8, alpha = 0.5, inherit.aes = FALSE) +
  
  # Violin plot and all points from summary_AF
  geom_violin(aes(fill = Region), alpha = 0.3) +
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1) +
  
  # Points for highlighted SNPs
  geom_point(
    data = highlighted_data,
    aes(x = Region, y = allele_frequency, colour = Region),
    position = position_jitter(width = 0.1, height = 0),
    size = 1.5, inherit.aes = FALSE) +
  
  labs(x = "Region", y = "Allele Frequency") +
  scale_color_manual(values = pal[c(1, 3, 5)]) +
  scale_fill_manual(values = pal[c(1, 3, 5)]) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
af_plot
```

Statistics to see if AF is different between regions (KS test)
```{r}
# Separate the data by region
central_af <- summary_AF %>% filter(Region == "Central") %>% pull(allele_frequency)
hybrid_af <- summary_AF %>% filter(Region == "Hybrid") %>% pull(allele_frequency)
north_af <- summary_AF %>% filter(Region == "North") %>% pull(allele_frequency)

# Perform KS test between Central and Hybrid regions
ks_test_central_hybrid <- ks.test(central_af, hybrid_af)
ks_test_central_hybrid

# Perform KS test between Central and North regions
ks_test_central_north <- ks.test(central_af, north_af)
ks_test_central_north

# Perform KS test between Hybrid and North regions
ks_test_hybrid_north <- ks.test(hybrid_af, north_af)
ks_test_hybrid_north

#means of AF per region
AF_means <- summarySE(summary_AF, measurevar="allele_frequency",
                     groupvars=c("Region"), na.rm=TRUE)
```

Larvae PCA
```{r}
#fixed<- highlighted_data |> filter(Region == "Central") |> pull(SNP)
filtered_T0<- gl.keep.loc(ten_larvae_T0, loc.list = fixed)
#fixed<- highlighted_data |> filter(Region == "Central") |> pull(SNP)
filtered_T0<- gl.keep.pop(filtered_T0, pop.list = c("DavxDav_T0","PalxPal_T0","PalxDav_T0","JewxJew_T0","JewxDav_T0","JewxPal_T0"))

pcoa<- gl.pcoa(filtered_T0)
pca<- gl.pcoa.plot(pcoa, filtered_T0, scale = F)
explained_variance <- c(48.5, 5.8)  # % Variation for PC1 and PC2

pcascore<- as.data.frame(pcoa$scores) |> rownames_to_column(var = "Labels")
pops<- as.data.frame(filtered_T0$pop)|> dplyr::rename(pop = `filtered_T0$pop`)

pcascore<- cbind(pops, pcascore)

pcascore <- pcascore |>
  mutate(Region = case_when(
    pop %in% c("DavxDav_T0", "PalxDav_T0", "PalxPal_T0") ~ "Central",
    pop %in% c("JewxJew_T0", "ParxPar_T0", "SwixSwi_T0") ~ "North",
    pop %in% c("JewxDav_T0", "JewxPal_T0", "ParxDav_T0", "SwixDav_T0") ~ "Inter-region"
  ))

# Plot PCA loadings for PC1 and PC2
pca1<- ggplot(pcascore, aes(x = PC1, y = PC2, colour = Region)) +
  geom_point(size = 3) +  # Add points
  #geom_text(vjust = -0.5, hjust = 0.5) +  # Add labels near points
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Reef") +
  #coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  theme_Publication() +
  scale_colour_manual(values = pal[c(1, 3, 5)])
pca1
```

Adult PCA
```{r}
adult <- readRDS("/Users/alexmacadam/Documents/GitHub/AGF-DArT/Data/ten_adult_filter.RDS")
#Filter only used parents
adult <- gl.keep.ind(adult, ind.list=c(
'A.ten_Davies_1',
'A.ten_Davies_10',
'A.ten_Davies_13',
'A.ten_Davies_6',
'A.ten_Davies_8',
'A.ten_B.Jewell_2',
'A.ten_B.Jewell_3',
'A.ten_B.Jewell_6',
'A.ten_B.Jewell_8',
'A.ten_Palms_1',
'A.ten_Palms_2',
'A.ten_Palms_4',
'A.ten_Palms_5',
'A.ten_Parke_2',
'A.ten_Parke_4',
'A.ten_Parke_5',
'A.ten_Switzer_4',
'A.ten_Switzer_6',
'A.ten_Switzer_7'))

#all sig SNPs
adult2<- gl.keep.loc(adult, loc.list = sig_als)
```

```{r}
pcoa2<- gl.pcoa(adult2)
pca2<- gl.pcoa.plot(pcoa2, adult2, scale = F, ellipse = T, plevel = 0.95)
explained_variance <- c(41.6, 7.9)  # % Variation for PC1 and PC2

pcascore<- as.data.frame(pcoa2$scores) |> rownames_to_column(var = "Labels")
pops<- as.data.frame(adult2$pop)|> dplyr::rename(pop = `adult2$pop`)

pcascore<- cbind(pops, pcascore)

pcascore <- pcascore %>%
  mutate(Region = case_when(
    pop %in% c("Davies", "Palms") ~ "Central",
    pop %in% c("Jewell", "Parke", "Switzer") ~ "North",
    TRUE ~ NA_character_  # for any unmatched reefs
  ))

# Plot PCA loadings for PC1 and PC2
pca2<- ggplot(pcascore, aes(x = PC1, y = PC2, colour = Region, shape = pop)) +
  geom_point(size = 3) +  # Add points
  #geom_text(vjust = -0.5, hjust = 0.5) +  # Add labels near points
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Reef") +
  #coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  scale_colour_Publication()
pca2

#just Fixed SNPs
adult3<- gl.keep.loc(adult, loc.list = fixed)

pcoa3<- gl.pcoa(adult3)
pca3<- gl.pcoa.plot(pcoa3, adult3, scale = F, ellipse = F)
explained_variance <- c(49.5, 10.2)  # % Variation for PC1 and PC2

pcascore<- as.data.frame(pcoa3$scores) |> rownames_to_column(var = "Labels")
pops<- as.data.frame(adult3$pop)|> dplyr::rename(pop = `adult3$pop`)

pcascore<- cbind(pops, pcascore)

pcascore <- pcascore %>%
  mutate(Region = case_when(
    pop %in% c("Davies", "Palms") ~ "Central",
    pop %in% c("Jewell", "Parke", "Switzer") ~ "North",
    TRUE ~ NA_character_  # for any unmatched reefs
  ))

# Plot PCA loadings for PC1 and PC2
pca3<- ggplot(pcascore, aes(x = PC1, y = PC2, colour = Region, shape = pop)) +
  geom_point(size = 3) +  # Add points
  #geom_text(vjust = -0.5, hjust = 0.5) +  # Add labels near points
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Reef") +
  #coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) +
  theme_Publication() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  scale_colour_Publication()
pca3

pca2 + pca3

library(vegan)
pc_mat <- pcascore[, c("PC1","PC2","PC3","PC4","PC5")]
dist_mat <- dist(pc_mat, method = "euclidean")
adonis2(dist_mat ~ Region, data = pcascore)
adonis2(dist_mat ~ pop, data = pcascore)
```

```{r}
#coloured by deviances
deviances<- read.csv("/Users/alexmacadam/Documents/GitHub/AGF-DArT/data/pm_deviance.csv")
larv<- deviances |> filter(Lifestage == "larv")
larv_dam<- larv |> filter(group == "Dams") |> dplyr::rename(Labels = parent) |> dplyr::select(Labels, mean)
larv_sire<- larv |> filter(group == "Sires") |> dplyr::rename(Labels = parent) |> dplyr::select(Labels, mean)

library(stringr)
pcascore$Labels <- str_replace(pcascore$Labels, ".*n_", "")
pcascore$Labels <- str_replace_all(pcascore$Labels, "_", "")
pcascore$Labels <- str_replace_all(pcascore$Labels, "B.", "")

pcascore3<- left_join(pcascore, larv_dam, by = "Labels")

pca4<- ggplot(pcascore3, aes(x = PC1, y = PC2)) +
  geom_point(aes(colour = mean), size=3) + 
  geom_point(shape = 1, size = 3,colour = "black") +
  #geom_text(vjust = -0.5, hjust = 0.5) +
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Maternal posterior means") +
  theme_Publication() +
  scale_color_gradient2(low = "blue", high = "red", space = "Lab")
  #coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  #scale_colour_Publication()
pca4

pcascore5<- left_join(pcascore, larv_sire, by = "Labels")

pca5<- ggplot(pcascore4, aes(x = PC1, y = PC2)) +
  geom_point(aes(colour = mean), size=3) + 
  geom_point(shape = 1, size = 3,colour = "black") +
  #geom_text(vjust = -0.5, hjust = 0.5) +
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Paternal posterior means") +
  theme_Publication() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  scale_color_gradient2(mid = "white",low = "blue", high = "red", space = "Lab")
  #coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  #scale_colour_Publication() 
pca5
```

```{r}
(pca1 + pca2) / (pca3 + pca4)
```

```{r}
adult_list<- c('A.ten_Davies_1',
'A.ten_Davies_10',
'A.ten_Davies_13',
'A.ten_Davies_6',
'A.ten_Davies_8',
'A.ten_B.Jewell_2',
'A.ten_B.Jewell_3',
'A.ten_B.Jewell_6',
'A.ten_B.Jewell_8',
'A.ten_Palms_1',
'A.ten_Palms_2',
'A.ten_Palms_4',
'A.ten_Palms_5',
'A.ten_Parke_2',
'A.ten_Parke_4',
'A.ten_Parke_5',
'A.ten_Switzer_4',
'A.ten_Switzer_6',
'A.ten_Switzer_7')

b<- adult$other$loc.metrics$maf

adult2
```

```{r}
#load parent posterior means
post_means_larv<- read.csv("/Users/alexmacadam/Documents/GitHub/AGF-DArT/data/post_means_larv.csv") |>
  dplyr::select(Sex, Region, IID, mean, se)

#make SNP matrix
parent_af<- as.data.frame(adult2) |> t()

#rename samples to match larvae
colnames(parent_af) <- colnames(parent_af) %>%
  str_replace_all(c(
    "A.ten_Davies_" = "Davies",
    "A.ten_B.Jewell_" = "Jewell",
    "A.ten_Palms_" = "Palms",
    "A.ten_Parke_" = "Parke",
    "A.ten_Switzer_" = "Switzer"
  ))
```

```{r}
library(tidyverse)
library(ggrepel)

# 1. Extract significant SNPs and their effects
sig_idx <- which(p.adjust(cor1$p, method = "BH") < 0.05)
sig_snps <- colnames(alf2)[1:13057][sig_idx]
sig_effects <- cor1$r[sig_idx]

# 2. Calculate weighted allele scores for each parent
weighted_scores <- parent_af * sig_effects # Weight by correlation coefficient
parent_scores <- colMeans(weighted_scores, na.rm = TRUE) %>% 
  data.frame(IID = names(.), Score = .)

# 3. Merge with posterior means
plot_data <- post_means_larv %>%
  inner_join(parent_scores, by = "IID") %>%
  mutate(ParentType = ifelse(Sex == "Dams", "Dams", "Sires"))

# 4. Plot data
plot_data |>
  ggplot(aes(x = Score, y = mean, color = Region)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  facet_wrap(~Sex)
```

```{r}
# 1. Extract significant SNPs (absolute values for effects)
sig_idx <- which(p.adjust(cor1$p, method = "BH") < 0.05)
sig_snps <- colnames(alf2)[1:13057][sig_idx]

# 2. Calculate unweighted positive allele scores
# Use absolute values to make all effects positive
parent_scores <- colMeans(parent_af, na.rm = TRUE) %>% 
  data.frame(IID = names(.), Score = .)

# 3. Merge with posterior means
plot_data <- post_means_larv %>%
  inner_join(parent_scores, by = "IID") %>%
  mutate(Sex = factor(Sex, levels = c("Dams", "Sires")))

# 4. Plot data
final_plot <- plot_data |>
  ggplot(aes(x = Score, y = mean, color = Region)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  facet_wrap(~Sex) +
  labs(x = "Average allele frequency (all positive effects)",
       y = "Posterior mean deviance",
       title = "All SNPs treated as having positive effects") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5))

# Add regression stats
add_stats <- function(plot, data) {
  model <- lm(mean ~ Score, data = data)
  r2 <- format(summary(model)$r.squared, digits = 3)
  p <- format(anova(model)$'Pr(>F)'[1], digits = 3)
  
  plot + 
    annotate("text",
             x = min(data$Score, na.rm = TRUE),
             y = max(data$mean, na.rm = TRUE),
             label = paste0("R² = ", r2, "\np = ", p),
             hjust = 0, vjust = 1)
}

# Apply to each sex separately
final_plot <- final_plot +
  geom_text(
    data = plot_data %>% group_by(Sex) %>%
      do({
        model <- lm(mean ~ Score, data = .)
        data.frame(
          Sex = first(.$Sex),
          x = min(.$Score, na.rm = TRUE),
          y = max(.$mean, na.rm = TRUE),
          label = paste0("R² = ", format(summary(model)$r.squared, digits = 3),
                       "\np = ", format(anova(model)$'Pr(>F)'[1], digits = 3))
        )
      }),
    aes(x = x, y = y, label = label),
    hjust = 0, vjust = 1,
    inherit.aes = FALSE
  )

final_plot
```

########
Genotype of sig SNPs in parents vs survival deviance
```{r}
#Extract only significant SNPs 0.05
sig_list_05 <- highly_pred_SNPs |> 
  rownames_to_column(var = "SNP") |> 
  dplyr::select(SNP, r, p)

sig_list_05_pos<- sig_list_05 |>
  dplyr::filter(r > 0)

sig_list_05_neg<- sig_list_05 |>
  dplyr::filter(r < 0)
```

```{r}
#load adult data
ten_adult <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_adult_filter.RDS")

# Filter out only significant SNPs

#0.05 neg
ten_adult2_neg<- gl.keep.loc(ten_adult, loc.list=c('9498501-47-C/T','26745102-51-C/A','15453551-57-C/T','26738349-68-T/G','11471890-18-C/A','11466799-26-T/G','50826019-43-C/A','26723097-63-C/A','15475296-11-C/T','11490533-61-C/T','17479105-58-A/G','11472298-35-T/C','17483183-14-A/T','11461901-9-T/C','15474193-24-A/G','11499615-11-A/T','15475949-14-C/A','17481402-32-A/T'))

#0.05 pos
ten_adult2_pos<- gl.keep.loc(ten_adult, loc.list=c('15457567-60-T/C','15478027-42-A/T','26735296-20-T/C','15470469-27-T/G','15464910-62-C/T','15476258-41-A/G','15471002-32-A/G','11482155-24-T/C','15477968-19-C/A','9489413-39-A/G','15466554-20-G/A','11472783-25-A/G','15471368-21-G/T','11499678-10-T/C'))
```

```{r}
#relatedness of parent alleles
#join 
sig_adult<- gl.join(ten_adult2_pos, ten_adult2_neg)

#filter only parents
sig_adult<- gl.keep.ind(sig_adult, ind.list = c("A.ten_B.Jewell_6", "A.ten_B.Jewell_3", "A.ten_B.Jewell_8", "A.ten_B.Jewell_7", "A.ten_B.Jewell_2", "A.ten_B.Jewell_5", "A.ten_Parke_2", "A.ten_Parke_5", " A.ten_Parke_4", "A.ten_Switzer_6", "A.ten_Switzer_7", "A.ten_Switzer_2", "A.ten_Switzer_4", "A.ten_Palms_1", "A.ten_Palms_4", "A.ten_Palms_2", "A.ten_Palms_5", "A.ten_Davies_13", "A.ten_Davies_8", "A.ten_Davies_10", "A.ten_Davies_1", "A.ten_Davies_6"))

#grm
grm<- gl.grm(sig_adult)
grm_plot<- gl.plot.heatmap(grm)

dist<- dist(grm)

pcoa<- gl.pcoa(sig_adult)
pca<- gl.pcoa.plot(pcoa, sig_adult)

explained_variance <- c(39.2, 13.9)  # % Variation for PC1 and PC2

pcascore<- as.data.frame(pcoa$scores) |> rownames_to_column(var = "Labels")
pops<- as.data.frame(sig_adult$pop)|> dplyr::rename(pop = `sig_adult$pop`)

pcascore<- cbind(pops, pcascore)

# Plot PCA loadings for PC1 and PC2
pca1<- ggplot(pcascore, aes(x = PC1, y = PC2, colour = pop)) +
  geom_point(size = 3) +  # Add points
  #geom_text(vjust = -0.5, hjust = 0.5) +  # Add labels near points
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Reef") +
  coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  theme_Publication() +
  scale_colour_Publication()
pca1

```

##################
```{r}
#load adult significant SNP allele frequency and survival deviance data
#ad_alf plot is 0.10 sig
ad_alf<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/alf_corr_plot.csv")
#ad_alf plot 2 is 0.05 sig
ad_alf2<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/alf_corr_plot2.csv")
#Deviance survival (parental contribution)
cov <- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/aten_adult.PRS2.cov.csv") |> dplyr::select(IID, Deviance)
```

```{r}
#plot 0.10 sig
ad_alf |>
  ggplot() +
  geom_boxplot(aes(x = Reef, y = alf)) +
  theme_Publication() +
  scale_colour_Publication() +
  labs(y = "Allele Frequency")

#plot 0.05 sig
ad_alf2 |>
  ggplot() +
  geom_boxplot(aes(x = Reef, y = alf)) +
  theme_Publication() +
  scale_colour_Publication() +
  labs(y = "Allele Frequency")
```

```{r}
#join the covariate (parental contribution) with allele frequecies of Sig SNPs
#0.10 sig
alf3<- left_join(cov, ad_alf, by = "IID") |> na.omit()
#0.05 sig
alf4<- left_join(cov, ad_alf2, by = "IID") |> na.omit()
```

```{r}
#calcualte corr coefficient for the survival and Allele freq of significant SNPs (minus outlier A.ten_B.Jewell_7)
alf5<- alf3 |> filter(IID != "A.ten_B.Jewell_7")
cor_test_result1 <- cor.test(alf5$Deviance, alf5$alf)
cor_test_result2 <- cor.test(alf4$Deviance, alf4$alf)

# View the correlation coefficient and p-value
ad_cor<- cor_test_result1$estimate  # Correlation coefficient
ad_p<- cor_test_result1$p.value   # P-value

#adults boxplots by reef
alf_point<- alf3 |> filter(IID != "A.ten_B.Jewell_7") |>
  ggplot() +
    geom_point(aes(x=alf, y=Deviance, color=Reef), size = 2) +
    geom_point(aes(x = 0.17876344, y = 2.67747089), color = "#fdb462", size = 2, shape = 21, fill = NA) + #add the outlier in
    #geom_line(stat = "identity") +
    geom_smooth(aes(x=alf, y=Deviance), method=lm, se=FALSE, colour = "black") +
    annotate("text", x=0.17, y=2.4, label= "R = 0.658, p = 0.0008", size = 6) +
    theme_Publication() +
    #scale_color_manual(values=pal[c(1,2,3,4,5)]) +
    scale_colour_Publication() +
    #ylim(0.11,0.13) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  labs(x = "Allele Frequency", y = "Parental ICC")
alf_point
```
