---
title: "Untitled"
output: html_document
date: "2024-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}
# Install the latest version of dartR (v2)
#install.packages("dartR")
library(dartR)
library(tidyverse)
library(ggpubr)
library(adegenet)
library(poppr)
library(hierfstat)
library(reshape2)
library(RColorBrewer)
library(scales)
library(LEA)
library(wesanderson)
```

```{r}
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

pal <- wes_palette("Zissou1", 5, type = "discrete")
```

```{r}
# Function for summarising standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

set working directory
```{r, echo = FALSE, comment= FALSE, message=FALSE}
#setwd("/Users/alexmacadam/Desktop/A_tenuis/larvae/allele_freq/assoc_files")
```

Fisher model
#####
```{r}
ten_larvae <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_larvae_filter.RDS")
#keep T0
ten_larvae_amb<- gl.keep.pop(ten_larvae, pop.list = c('Amb'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_hot<- gl.keep.pop(ten_larvae, pop.list = c('Hot'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)

#allele freq central amb
amb_cen<- gl.keep.pop(ten_larvae_amb, pop.list= c("pure_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_cen<- gl.alf(amb_cen) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq hybrid amb
amb_hyb<- gl.keep.pop(ten_larvae_amb, pop.list= c("hybrid_north_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_hyb<- gl.alf(amb_hyb) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq north amb
amb_nor<- gl.keep.pop(ten_larvae_amb, pop.list= c("pure_north"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_nor<- gl.alf(amb_nor) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq central hot
hot_cen<- gl.keep.pop(ten_larvae_hot, pop.list= c("pure_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_cen<- gl.alf(hot_cen) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#allele freq hybrid hot
hot_hyb<- gl.keep.pop(ten_larvae_hot, pop.list= c("hybrid_north_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_hyb<- gl.alf(hot_hyb) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#allele freq north hot
hot_nor<- gl.keep.pop(ten_larvae_hot, pop.list= c("pure_north"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_nor<- gl.alf(hot_nor) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#Join ambient and hot datat for each region cross
af_cen<- left_join(af_amb_cen, af_hot_cen, by = "SNP") |> mutate(Region = "Central")
af_hyb<- left_join(af_amb_hyb, af_hot_hyb, by = "SNP") |> mutate(Region = "Hybrid")
af_nor<- left_join(af_amb_nor, af_hot_nor, by = "SNP") |> mutate(Region = "North")

#combine all data
af<- rbind(af_cen, af_hyb, af_nor)
```

Central
```{r}
# Load necessary library
library(dplyr)

# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_cen2<- af_cen %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_cen2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_cen2)) {
    # Extract counts for each SNP
    count1_allele <- af_cen2$alf1_amb[i]
    count1_non_allele <- af_cen2$alf2_amb[i]
    count2_allele <- af_cen2$alf1_hot[i]
    count2_non_allele <- af_cen2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_cen2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_cen2$p_adj <- p.adjust(af_cen2$p_value, method = "BH")
```

Hybrid
```{r}
# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_hyb2<- af_hyb %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_hyb2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_hyb2)) {
    # Extract counts for each SNP
    count1_allele <- af_hyb2$alf1_amb[i]
    count1_non_allele <- af_hyb2$alf2_amb[i]
    count2_allele <- af_hyb2$alf1_hot[i]
    count2_non_allele <- af_hyb2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_hyb2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_hyb2$p_adj <- p.adjust(af_hyb2$p_value, method = "BH")
```

North
```{r}
# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_nor2<- af_nor %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_nor2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_nor2)) {
    # Extract counts for each SNP
    count1_allele <- af_nor2$alf1_amb[i]
    count1_non_allele <- af_nor2$alf2_amb[i]
    count2_allele <- af_nor2$alf1_hot[i]
    count2_non_allele <- af_nor2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_nor2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_nor2$p_adj <- p.adjust(af_nor2$p_value, method = "BH")
```

#extract the significant SNPs
```{r}
sig1<- af_cen2 |> filter(p_adj < 0.05) |> pull(SNP)
sig2<- af_hyb2 |> filter(p_adj < 0.05) |> pull(SNP)
sig<- c(sig1, sig2)
```

No filtering!
```{r}
af2_all<- af |>
  dplyr::select(SNP, Region, alf1_amb, alf1_hot) |> 
  dplyr::rename(amb = alf1_amb, hot = alf1_hot) |>
  pivot_longer(cols=c('amb', 'hot'),
                    names_to='temp',
                    values_to='alf')

# Assuming your original column is called 'alf'
af2_all$alf_new <- ifelse(af2_all$alf > 0.5, 1 - af2_all$alf, af2_all$alf)
```

```{r}
cols<- c("#277da1","#f94144")

af_plot_all<- af2_all |> 
  ggplot() +
  geom_density(aes(alf_new, fill = temp), adjust = 1.2, alpha = 0.5) +
  facet_wrap(~Region, nrow = 1) +
  theme_classic() +
  scale_fill_manual(values=cols) +
  labs(x="Allele Frequency", y="Density")
af_plot_all
```

Filtered for significant SNPs
```{r}
af2_sig<- af |> filter(SNP %in% sig) |>
  dplyr::select(SNP, Region, alf1_amb, alf1_hot) |> 
  dplyr::rename(amb = alf1_amb, hot = alf1_hot) |>
  pivot_longer(cols=c('amb', 'hot'),
                    names_to='temp',
                    values_to='alf')

# change AF to MAF
af2_sig$alf_new <- ifelse(af2_sig$alf > 0.5, 1 - af2_sig$alf, af2_sig$alf)
```

plot AF
```{r}
af_plot<- af2_sig |> 
  ggplot() +
  geom_density(aes(alf_new, fill = temp), adjust = 1, alpha = 0.5) +
  facet_wrap(~Region, nrow = 1) +
  theme_classic() +
  scale_fill_manual(values=cols) +
  labs(x="Allele Frequency", y="Density")
af_plot
```

No difference in the hot, well heat adapted. the hybirds follow this so may have inherited some beneficial heat associated alleles. central has a big difference in the hot treatments. Many allele freqs move to nearer zero. however, some went towards 0.5, which may indicate heterozygote advantage or heterosis. overdominance, where the heterozygous genotype has a higher adaptive value than either homozygous genotype.

KS test for distribution of significant SNPs
```{r}
#central
ks_central<- af2_sig |> filter(Region == "Central") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_central$amb, ks_central$hot)

#hybrid
ks_hybrid<- af2_sig |> filter(Region == "Hybrid") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_hybrid$amb, ks_hybrid$hot)

#north
ks_north<- af2_sig |> filter(Region == "North") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_north$amb, ks_north$hot)
```
