---
title: "05_A.ten_larvae_diversity"
output: html_document
date: "2024-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

alpha= within popualtion, eg heterozygosity, allelic richness, Nucleotide diversity
beta= between populations, eg, Fst, AMOVA, Jaccard Index, Bray curtis

#calculate alpha and beta diversity for q=0, q=1 and q=2
#within population diversity, i.e. alpha values:
#q=0 represents allelic richness index
#q=1 Shannon information index
#q=2 heterozygosity index
#indices are represented by H values. These H values can then be converted to a common scale of effective numbers, D.
#For between population comparisons of diversity, i.e. beta values:
#q=0 represents number of allelic types not shared between pops
#q=1 measures the chance of a type of individuals providing #info on the sampling locality
#q=2 measured the chance of two random individuals from different localities having different allelic types

##############
load packages
```{r}
set.seed(123)
library(dartRverse)
library(tidyverse)
library(adegenet)
library(poppr)
library(hierfstat)
library(reshape2)
library(RColorBrewer)
library(scales)
library(LEA)
library(wesanderson)
library(tibble)
library(purrr)
library(ggthemes)
library(vegan)
```

Load personal theme
```{r}
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

#pallets
pd <- position_dodge(width=0.2)
pal <- wes_palette("Zissou1", 5, type = "discrete")
scalesize <- function(x) sprintf("%.1f", x)
pal2 <- c("#386cb0","#7fc97f","#fdb462")
```

PERCENT SURVIVAL function
```{r}
# Function for summarising survival standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
  
  #Use the function like this:
  #summarySE(data, measurevar="Value", groupvars=c("Metric1","Metric2"), na.rm=TRUE)
}
```

load meta data
```{r}
#load metadata
meta<- read.csv("/Users/alexmacadam/Documents/GitHub/AGF-DArT/Data/aten_metadata.csv") |>
  filter(lifestage == "larvae")

#remove models and summarise
sum<- meta |> filter(Temp == "T0", !family %in% c("Mod1","Mod2", "Mod3", "Mod4","Mod5","Mod6")) |>
  group_by(pop, indcross) |>
  dplyr::summarise(n = n())
```

Load data
```{r}
#dart data
ten_larvae <- readRDS("/Users/alexmacadam/Documents/GitHub/AGF-DArT/Data/ten_larvae_filter.RDS")
ten_larvae <- gl.drop.pop(ten_larvae, pop.list=c('ParxDav_T0', 'ParxDav_Amb', 'ParxDav_Hot'),mono.rm=FALSE,recalc=TRUE)
```

Remove model samples
```{r}
ten_larvae<- gl.drop.ind(ten_larvae, ind.list = c("Mod1-JewBotxDav6Bul-Cn1-T0-rep1","Mod2-Par6BulxPar6Bul-Cn3-T0-rep3","Mod3-Par6BulxDav6Bul-Cn5-T0-rep5","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep1","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep3","Mod1-JewBotxDav6Bul-Cn1-T0-rep2","Mod2-Par6BulxPar6Bul-Cn3-T0-rep4","Mod3-Par6BulxDav6Bul-Cn5-T0-rep6","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep2","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep4","Mod1-JewBotxDav6Bul-Cn1-T0-rep3","Mod2-Par6BulxPar6Bul-Cn3-T0-rep5","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep1","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep3","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep5","Mod1-JewBotxDav6Bul-Cn1-T0-rep4","Mod2-Par6BulxPar6Bul-Cn3-T0-rep6","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep2","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep4","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep6","Mod1-JewBotxDav6Bul-Cn1-T0-rep5","Mod3-Par6BulxDav6Bul-Cn5-T0-rep1","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep3","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep5","Mod1-JewBotxDav6Bul-Cn1-T0-rep6","Mod3-Par6BulxDav6Bul-Cn5-T0-rep2","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep4","Mod5-Swi6BulxDav6Bul-Cn9-T0-rep6","Mod2-Par6BulxPar6Bul-Cn3-T0-rep1","Mod3-Par6BulxDav6Bul-Cn5-T0-rep3","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep5","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep1","Mod2-Par6BulxPar6Bul-Cn3-T0-rep2","Mod3-Par6BulxDav6Bul-Cn5-T0-rep4","Mod4-Swi6BulxSwi6Bul-Cn7-T0-rep6","Mod6-MunPalBulxDav6Bul-Cn13-T0-rep2","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep5","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep3","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep5","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep1","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep5","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep6","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep4","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep6","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep2","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep6","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep5","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep3","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep1","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep6","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep4","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep2","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep1","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep5","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep1","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep3","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep1","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep2","Mod3-Par6BulxDav6Bul-Bt137-T3-Amb-rep6","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep2","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep4","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep2","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep1","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep3","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep3","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep5","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep3","Mod1-JewBotxDav6Bul-Bt122-T3-Amb-rep2","Mod2-Par6BulxPar6Bul-Bt109-T3-Amb-rep4","Mod4-Swi6BulxSwi6Bul-Bt125-T3-Amb-rep4","Mod5-Swi6BulxDav6Bul-Bt122-T3-Amb-rep6","Mod6-MunPalBulxDav6Bul-Bt87-T3-Amb-rep4","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep3","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep1","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep3","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep2","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep4","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep2","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep4","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep3","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep5","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep3","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep5","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep4","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep2","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep6","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep4","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep6","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep5","Mod3-Par6BulxDav6Bul-Bt159-T2-Hot-rep3","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep5","Mod1-JewBotxDav6Bul-Bt121-T2-Hot-rep6","Mod5-Swi6BulxDav6Bul-Bt19-T2-Hot-rep6","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep1","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep1","Mod4-Swi6BulxSwi6Bul-Bt126-T2-Hot-rep2","Mod6-MunPalBulxDav6Bul-Bt136-T2-Hot-rep2"), mono.rm = FALSE, recalc=TRUE)
```

Create Pre, Hot and Ambient gl objects
```{r}
ten_larvae_Amb<- gl.keep.pop(ten_larvae, pop.list = c('Amb'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_Hot<- gl.keep.pop(ten_larvae, pop.list = c('Hot'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_T0<- gl.keep.pop(ten_larvae, pop.list = c('T0'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
```

#############
By reef
50 iterations of diversity metrics
Pre
```{r}
# Check how many samples per population
table(pop(ten_larvae_T0))

# Define the minimum sample size and number of draws
min_sample_size_T0 <- 12
num_draws_T0 <- 50

# Initialize a list to store results for each draw
subsample_results_T0 <- vector("list", num_draws_T0)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_T0) {
  subsample_results_T0[[i]] <- gl.subsample.ind(
    x = ten_larvae_T0,          # The genlight object
    n = min_sample_size_T0,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_T0[[1]]))

# Calculate diversity metrics for each draw
diversity_results_T0 <- lapply(subsample_results_T0, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_T0 <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_T0 <- diversity_results_T0[[i]]
  
  # Create a dataframe with the relevant columns
  df_T0 <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_T0$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_T0$zero_H_alpha_sd,
    one_H_alpha = iteration_data_T0$one_H_alpha,
    one_H_alpha_sd = iteration_data_T0$one_H_alpha_sd,
    two_H_alpha = iteration_data_T0$two_H_alpha,
    two_H_alpha_sd = iteration_data_T0$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_T0[[i]] <- df_T0
}

# Combine all the dataframes into one
diversity_combined_T0 <- do.call(rbind, diversity_summary_list_T0) |>
  rownames_to_column(var = "population") |>
  plyr::mutate(Reefcross = substr(population, 1, 7))|>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Reefcross, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_T0<- rbind(long_ar_T0, long_SI_T0, long_He_T0)

# Calculate SE
long_diversity_T0<- long_diversity_T0 |> mutate(se = sd/sqrt(48))

diversity_sum_T0 <- summarySE(long_diversity_T0, measurevar="Value",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::select(Metric, Reefcross, mean)

diversity_se_T0 <- summarySE(long_diversity_T0, measurevar="se",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_pre <- cbind(diversity_sum_T0, diversity_se_T0) |>
  mutate(Temp = "Pre")
```

Amb
```{r}
# Check how many samples per population
table(pop(ten_larvae_Amb))

# Define the minimum sample size and number of draws
min_sample_size_Amb <- 12
num_draws_Amb <- 50

# Initialize a list to store results for each draw
subsample_results_Amb <- vector("list", num_draws_Amb)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_Amb) {
  subsample_results_Amb[[i]] <- gl.subsample.ind(
    x = ten_larvae_Amb,          # The genlight object
    n = min_sample_size_Amb,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_Amb[[1]]))

# Calculate diversity metrics for each draw
diversity_results_Amb <- lapply(subsample_results_Amb, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_Amb <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_Amb <- diversity_results_Amb[[i]]
  
  # Create a dataframe with the relevant columns
  df_Amb <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_Amb$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_Amb$zero_H_alpha_sd,
    one_H_alpha = iteration_data_Amb$one_H_alpha,
    one_H_alpha_sd = iteration_data_Amb$one_H_alpha_sd,
    two_H_alpha = iteration_data_Amb$two_H_alpha,
    two_H_alpha_sd = iteration_data_Amb$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_Amb[[i]] <- df_Amb
}

# Combine all the dataframes into one
diversity_combined_Amb <- do.call(rbind, diversity_summary_list_Amb) |>
  rownames_to_column(var = "population") |>
  plyr::mutate(Reefcross = substr(population, 1, 7))|>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Reefcross, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Reefcross, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_Amb<- rbind(long_ar_Amb, long_SI_Amb, long_He_Amb)

# Calculate SE
long_diversity_Amb<- long_diversity_Amb |> mutate(se = sd/sqrt(48))

diversity_sum_Amb <- summarySE(long_diversity_Amb, measurevar="Value",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::select(Metric, Reefcross, mean)

diversity_se_Amb <- summarySE(long_diversity_Amb, measurevar="se",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_amb <- cbind(diversity_sum_Amb, diversity_se_Amb) |>
  mutate(Temp = "Amb")
```

Hot
```{r}
# Check how many samples per population
table(pop(ten_larvae_Hot))

# Define the minimum sample size and number of draws
min_sample_size_Hot <- 6
num_draws_Hot <- 50

# Initialize a list to store results for each draw
subsample_results_Hot <- vector("list", num_draws_Hot)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_Hot) {
  subsample_results_Hot[[i]] <- gl.subsample.ind(
    x = ten_larvae_Hot,          # The genlight object
    n = min_sample_size_Hot,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_Hot[[1]]))

# Calculate diversity metrics for each draw
diversity_results_Hot <- lapply(subsample_results_Hot, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_Hot <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_Hot <- diversity_results_Hot[[i]]
  
  # Create a dataframe with the relevant columns
  df_Hot <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_Hot$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_Hot$zero_H_alpha_sd,
    one_H_alpha = iteration_data_Hot$one_H_alpha,
    one_H_alpha_sd = iteration_data_Hot$one_H_alpha_sd,
    two_H_alpha = iteration_data_Hot$two_H_alpha,
    two_H_alpha_sd = iteration_data_Hot$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_Hot[[i]] <- df_Hot
}

# Combine all the dataframes into one
diversity_combined_Hot <- do.call(rbind, diversity_summary_list_Hot) |>
  rownames_to_column(var = "population") |>
  plyr::mutate(Reefcross = substr(population, 1, 7)) |>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Reefcross, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Reefcross, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Reefcross, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Reefcross, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_Hot<- rbind(long_ar_Hot, long_SI_Hot, long_He_Hot)

# Calculate SE
long_diversity_Hot<- long_diversity_Hot |> mutate(se = sd/sqrt(48))

diversity_sum_Hot <- summarySE(long_diversity_Hot, measurevar="Value",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::select(Metric, Reefcross, mean)

diversity_se_Hot <- summarySE(long_diversity_Hot, measurevar="se",
                     groupvars=c("Metric","Reefcross"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_hot <- cbind(diversity_sum_Hot, diversity_se_Hot) |>
  mutate(Temp = "Hot")
```

```{r}
#join dataframes and add Region crosses column
diversity_summary <- rbind(diversity_pre, diversity_amb, diversity_hot) |>
    mutate(Region = case_when(
    grepl("DavxDav|PalxPal|PalxDav", Reefcross) ~ "Central",
    grepl("JewxDav|JewxPal|SwixDav", Reefcross) ~ "Hybrid",
    grepl("JewxJew|ParxPar|SwixSwi", Reefcross) ~ "North",
    TRUE ~ NA_character_  # Assign NA if none of the conditions match
  )) |>
  mutate(Temp = factor(Temp, levels = c("Pre","Amb","Hot")))
```

```{r}
#saveRDS(diversity_summary, file = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/diversity_metrics/diversity_summary.rds")
diversity_summary <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/diversity_metrics/diversity_summary.rds") |>
  mutate(Metric = factor(Metric, levels = c("Allelic Richness","Shannon Index","He")))

#create diversity by region
div_sum_sum <- summarySE(diversity_summary, measurevar="mean",
                     groupvars=c("Temp","Metric","Region"), na.rm=TRUE) |>
  dplyr::select(Temp, Metric, Region, mean, sd)
div_sum_se <- summarySE(diversity_summary, measurevar="SE",
                     groupvars=c("Temp","Metric","Region"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |> dplyr::select(Temp, Metric, Region, SE)

diversity_region <- left_join(div_sum_sum, div_sum_se, by = c("Temp", "Metric", "Region"))
```

```{r}
div_plot1<- diversity_summary |>
  ggplot(aes(x=Reefcross, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Region), position=position_dodge(width=0.8))+
  facet_grid(Metric~Temp, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
div_plot1

div_plot2<- diversity_summary |>
  ggplot(aes(x=Reefcross, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
div_plot2

div_plot3<- diversity_region |>
  ggplot(aes(x=Region, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  #scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
div_plot3
```

PERMANOVA Pre
```{r}
#permanova pre
div<- diversity_summary |>
  dplyr::filter(Temp == "Pre") |>
  dplyr::select(Reefcross, Region, Metric, mean) |>
  pivot_wider(names_from = Metric, values_from = mean)

matrix<- as.matrix(div[,3:5]) # response variables
div$regioncross<-as.factor(div$Region)

#manova<-manova(matrix~as.factor(Region), data=div)
#summary(manova)
#summary.aov(manova)

dist<-vegdist(matrix, method='bray')
dist

set.seed(123) #reproducible results
bird.div<-adonis2(dist~as.factor(regioncross), data=div, permutations=9999)
bird.div

birdMDS<-metaMDS(matrix, distance="bray", k=2, trymax=35, autotransform=TRUE) ##k is the number of dimensions
stressplot(birdMDS)
NMDS1<-birdMDS$points[,1]
NMDS2<-birdMDS$points[,2]
bird.plot<- cbind(div,NMDS1, NMDS2)
# remove duplicated columns
b.plot <- bird.plot[,!duplicated(colnames(bird.plot))]

p<-ggplot(b.plot, aes(NMDS1, NMDS2, color=regioncross))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  geom_text(data=b.plot,aes(NMDS1, NMDS2, label=regioncross), position=position_jitter(.2))+
  theme_minimal()
p
```

PERMANOVA amb vs hot
```{r}
div2<- diversity_summary |>
  dplyr::filter(Temp %in% c("Amb","Hot")) |>
  dplyr::select(Reefcross, Region, Temp, Metric, mean) |>
  pivot_wider(names_from = Metric, values_from = mean)

matrix2<- as.matrix(div2[,4:6]) # response variables
div2$Temp<-as.factor(div2$Temp)
div2$regioncross<-as.factor(div2$Region)

#manova<-manova(matrix2~as.factor(Temp) * as.factor(Region), data=div2)
#summary(manova)
#summary.aov(manova)

dist<-vegdist(matrix2, method='bray')
dist

set.seed(123) #reproducible results
bird.div<-adonis2(dist~as.factor(Temp) * as.factor(regioncross), data=div2, permutations=9999)
bird.div

birdMDS<-metaMDS(matrix2, distance="bray", k=2, trymax=35, autotransform=TRUE) ##k is the number of dimensions
stressplot(birdMDS)
NMDS1<-birdMDS$points[,1]
NMDS2<-birdMDS$points[,2]
bird.plot<- cbind(div2,NMDS1, NMDS2)
# remove duplicated columns
b.plot <- bird.plot[,!duplicated(colnames(bird.plot))]

b.plot$region_temp <- paste(b.plot$regioncross, b.plot$Temp, sep = "_")

p<-ggplot(b.plot, aes(NMDS1, NMDS2, color=regioncross))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  geom_text(data=b.plot,aes(NMDS1, NMDS2, label=Temp), position=position_jitter(.2))+
  theme_minimal()
p
```

Kruskall wallis comparing deviance of region crosses and temp treatments
```{r}
diversity_summary_ah <- diversity_summary |> dplyr::filter(Temp %in% c("Amb","Hot"))

# Subset the data for each region cross
central_data <- filter(diversity_summary_ah, Region == "Central")
hybrid_data <- filter(diversity_summary_ah, Region == "Hybrid")
north_data <- filter(diversity_summary_ah, Region == "North")

# Kruskal-Wallis test for each metric (separately)

# Central region
# Allelic Richness
central_allelic_kw <- kruskal.test(mean ~ Temp, data = filter(central_data, Metric == "Allelic Richness"))
# Shannon Diversity
central_shannon_kw <- kruskal.test(mean ~ Temp, data = filter(central_data, Metric == "Shannon Index"))
# Expected Heterozygosity
central_heterozygosity_kw <- kruskal.test(mean ~ Temp, data = filter(central_data, Metric == "He"))

# Hybrid region
# Allelic Richness
hybrid_allelic_kw <- kruskal.test(mean ~ Temp, data = filter(hybrid_data, Metric == "Allelic Richness"))
# Shannon Diversity
hybrid_shannon_kw <- kruskal.test(mean ~ Temp, data = filter(hybrid_data, Metric == "Shannon Index"))
# Expected Heterozygosity
hybrid_heterozygosity_kw <- kruskal.test(mean ~ Temp, data = filter(hybrid_data, Metric == "He"))

# Northern region
# Allelic Richness
north_allelic_kw <- kruskal.test(mean ~ Temp, data = filter(north_data, Metric == "Allelic Richness"))
# Shannon Diversity
north_shannon_kw <- kruskal.test(mean ~ Temp, data = filter(north_data, Metric == "Shannon Index"))
# Expected Heterozygosity
north_heterozygosity_kw <- kruskal.test(mean ~ Temp, data = filter(north_data, Metric == "He"))

# Print the results for each test
central_allelic_kw
central_shannon_kw
central_heterozygosity_kw

hybrid_allelic_kw
hybrid_shannon_kw
hybrid_heterozygosity_kw

north_allelic_kw
north_shannon_kw
north_heterozygosity_kw
```

```{r}
# Region cross differences between Ambient and Hot

# Subset the data for each region cross
central_data <- filter(diversity_region, Region == "Central")
hybrid_data <- filter(diversity_region, Region == "Hybrid")
north_data <- filter(diversity_region, Region == "North")

# Reshape data to be wider
# Central
central_data2 <- central_data %>%
  pivot_wider(
    names_from = Temp,                   # Separate columns for 'Amb' and 'Hot'
    values_from = c(mean, sd, SE),           # Fill mean and SE values
    names_glue = "{Temp}_{.value}"       # Create descriptive column names
  )
# Hybrid
hybrid_data2 <- hybrid_data %>%
  pivot_wider(
    names_from = Temp,                   # Separate columns for 'Amb' and 'Hot'
    values_from = c(mean, sd, SE),           # Fill mean and SE values
    names_glue = "{Temp}_{.value}"       # Create descriptive column names
  )
# North
north_data2 <- north_data %>%
  pivot_wider(
    names_from = Temp,                   # Separate columns for 'Amb' and 'Hot'
    values_from = c(mean, sd, SE),           # Fill mean and SE values
    names_glue = "{Temp}_{.value}"       # Create descriptive column names
  )

# Make a difference column
# Central
central_data2 <- central_data2 |>
  mutate(mean_dif = Amb_mean - Hot_mean)
# Hybrid
hybrid_data2 <- hybrid_data2 |>
  mutate(mean_dif = Amb_mean - Hot_mean)
# North
north_data2 <- north_data2 |>
  mutate(mean_dif = Amb_mean - Hot_mean)

# Bind together
div_data <- rbind(central_data2, hybrid_data2, north_data2)

# Recalculate the SE
div_data$se_dif <- sqrt((div_data$Amb_sd^2 / 3) + (div_data$Hot_sd^2 / 3))
```

#################
#################
Population level
```{r}
popNames(ten_larvae)
ten_larvae2<- gl.merge.pop(ten_larvae, old = c('DavxDav_Amb','PalxPal_Amb','PalxDav_Amb'), new = 'Central_Amb')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('DavxDav_Hot','PalxPal_Hot','PalxDav_Hot'), new = 'Central_Hot')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('DavxDav_T0','PalxPal_T0','PalxDav_T0'), new = 'Central_T0')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixSwi_Amb','ParxPar_Amb','JewxJew_Amb'), new = 'North_Amb')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixSwi_Hot','ParxPar_Hot','JewxJew_Hot'), new = 'North_Hot')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixSwi_T0','ParxPar_T0','JewxJew_T0'), new = 'North_T0')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixDav_Amb','JewxDav_Amb','JewxPal_Amb'), new = 'Hybrid_Amb')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixDav_Hot','JewxDav_Hot','JewxPal_Hot'), new = 'Hybrid_Hot')
ten_larvae2<- gl.merge.pop(ten_larvae2, old = c('SwixDav_T0','JewxDav_T0','JewxPal_T0'), new = 'Hybrid_T0')
#ten_larvae<- gl.drop.pop(ten_larvae2, pop.list = c("ParxDav_Amb", "ParxDav_Hot", "ParxDav_T0"))
ten_larvae<- ten_larvae2
popNames(ten_larvae)
```

Create Pre, Hot and Ambient gl objects
```{r}
ten_larvae_Amb<- gl.keep.pop(ten_larvae, pop.list = c('Amb'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_Hot<- gl.keep.pop(ten_larvae, pop.list = c('Hot'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_T0<- gl.keep.pop(ten_larvae, pop.list = c('T0'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
```

#############
By region
50 iterations of diversity metrics
Pre
```{r}
# Check how many samples per population
table(pop(ten_larvae_T0))

# Define the minimum sample size and number of draws
min_sample_size_T0 <- 48
num_draws_T0 <- 50

# Initialize a list to store results for each draw
subsample_results_T0 <- vector("list", num_draws_T0)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_T0) {
  subsample_results_T0[[i]] <- gl.subsample.ind(
    x = ten_larvae_T0,          # The genlight object
    n = min_sample_size_T0,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_T0[[1]]))

# Calculate diversity metrics for each draw
diversity_results_T0 <- lapply(subsample_results_T0, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_T0 <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_T0 <- diversity_results_T0[[i]]
  
  # Create a dataframe with the relevant columns
  df_T0 <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_T0$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_T0$zero_H_alpha_sd,
    one_H_alpha = iteration_data_T0$one_H_alpha,
    one_H_alpha_sd = iteration_data_T0$one_H_alpha_sd,
    two_H_alpha = iteration_data_T0$two_H_alpha,
    two_H_alpha_sd = iteration_data_T0$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_T0[[i]] <- df_T0
}

# Combine all the dataframes into one
diversity_combined_T0 <- do.call(rbind, diversity_summary_list_T0) |>
  rownames_to_column(var = "Population") |>
  mutate(Region = case_when(
    grepl("Central", Population) ~ "Central",
    grepl("Hybrid", Population) ~ "Hybrid",
    grepl("North", Population) ~ "North",
    TRUE ~ NA_character_  # Assign NA if none of the conditions match
  )) |>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Region, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Region, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Region, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_T0 <- diversity_combined_T0 |>
  dplyr::select(Iteration, Region, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_T0<- rbind(long_ar_T0, long_SI_T0, long_He_T0)

# Calculate SE
long_diversity_T0<- long_diversity_T0 |> mutate(se = sd/sqrt(48))

diversity_sum_T0 <- summarySE(long_diversity_T0, measurevar="Value",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::select(Metric, Region, mean)

diversity_se_T0 <- summarySE(long_diversity_T0, measurevar="se",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_pre <- cbind(diversity_sum_T0, diversity_se_T0) |>
  mutate(Temp = "Pre")
```

Amb
```{r}
# Check how many samples per population
table(pop(ten_larvae_Amb))

# Define the minimum sample size and number of draws
min_sample_size_Amb <- 48
num_draws_Amb <- 50

# Initialize a list to store results for each draw
subsample_results_Amb <- vector("list", num_draws_Amb)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_Amb) {
  subsample_results_Amb[[i]] <- gl.subsample.ind(
    x = ten_larvae_Amb,          # The genlight object
    n = min_sample_size_Amb,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_Amb[[1]]))

# Calculate diversity metrics for each draw
diversity_results_Amb <- lapply(subsample_results_Amb, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_Amb <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_Amb <- diversity_results_Amb[[i]]
  
  # Create a dataframe with the relevant columns
  df_Amb <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_Amb$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_Amb$zero_H_alpha_sd,
    one_H_alpha = iteration_data_Amb$one_H_alpha,
    one_H_alpha_sd = iteration_data_Amb$one_H_alpha_sd,
    two_H_alpha = iteration_data_Amb$two_H_alpha,
    two_H_alpha_sd = iteration_data_Amb$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_Amb[[i]] <- df_Amb
}

# Combine all the dataframes into one
diversity_combined_Amb <- do.call(rbind, diversity_summary_list_Amb) |>
  rownames_to_column(var = "Population") |>
  mutate(Region = case_when(
    grepl("Central", Population) ~ "Central",
    grepl("Hybrid", Population) ~ "Hybrid",
    grepl("North", Population) ~ "North",
    TRUE ~ NA_character_  # Assign NA if none of the conditions match
  )) |>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Region, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Region, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Region, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_Amb <- diversity_combined_Amb |>
  dplyr::select(Iteration, Region, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_Amb<- rbind(long_ar_Amb, long_SI_Amb, long_He_Amb)

# Calculate SE
long_diversity_Amb<- long_diversity_Amb |> mutate(se = sd/sqrt(48))

diversity_sum_Amb <- summarySE(long_diversity_Amb, measurevar="Value",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::select(Metric, Region, mean)

diversity_se_Amb <- summarySE(long_diversity_Amb, measurevar="se",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_amb <- cbind(diversity_sum_Amb, diversity_se_Amb) |>
  mutate(Temp = "Amb")
```

Hot
```{r}
# Check how many samples per population
table(pop(ten_larvae_Hot))

# Define the minimum sample size and number of draws
min_sample_size_Hot <- 27
num_draws_Hot <- 50

# Initialize a list to store results for each draw
subsample_results_Hot <- vector("list", num_draws_Hot)

# Perform 50 random draws
set.seed(123) # For reproducibility
for (i in 1:num_draws_Hot) {
  subsample_results_Hot[[i]] <- gl.subsample.ind(
    x = ten_larvae_Hot,          # The genlight object
    n = min_sample_size_Hot,        # Number of individuals per subsample
    replace = FALSE,            # Sampling without replacement
    by.pop = TRUE,              # Subsample by population
    error.check = TRUE,         # Enable error checking
    verbose = 0                 # Reduce verbosity
  )
}

# Verify subsampled populations for the first draw
print("Subsampled populations for first draw:")
print(pop(subsample_results_Hot[[1]]))

# Calculate diversity metrics for each draw
diversity_results_Hot <- lapply(subsample_results_Hot, function(subsample) {
  gl.report.diversity(
    x = subsample,             # Subsampled genlight object
    verbose = 0                # Reduce verbosity
  )
})

# Initialize an empty list to store dataframes for each iteration
diversity_summary_list_Hot <- list()

# Loop through each iteration and extract the required data
for (i in 1:50) {
  # Extract data for the current iteration
  iteration_data_Hot <- diversity_results_Hot[[i]]
  
  # Create a dataframe with the relevant columns
  df_Hot <- data.frame(
    #Population = rownames(iteration_data$zero_H_alpha),  # Assuming populations are row names
    zero_H_alpha = iteration_data_Hot$zero_H_alpha,
    zero_H_alpha_sd = iteration_data_Hot$zero_H_alpha_sd,
    one_H_alpha = iteration_data_Hot$one_H_alpha,
    one_H_alpha_sd = iteration_data_Hot$one_H_alpha_sd,
    two_H_alpha = iteration_data_Hot$two_H_alpha,
    two_H_alpha_sd = iteration_data_Hot$two_H_alpha_sd,
    Iteration = paste0("iteration_", i)
  )
  
  # Append the dataframe to the list
  diversity_summary_list_Hot[[i]] <- df_Hot
}

# Combine all the dataframes into one
diversity_combined_Hot <- do.call(rbind, diversity_summary_list_Hot) |>
  rownames_to_column(var = "Population") |>
  mutate(Region = case_when(
    grepl("Central", Population) ~ "Central",
    grepl("Hybrid", Population) ~ "Hybrid",
    grepl("North", Population) ~ "North",
    TRUE ~ NA_character_  # Assign NA if none of the conditions match
  )) |>
  dplyr::rename(`Allelic Richness` = zero_H_alpha, `AR_sd` = zero_H_alpha_sd, `Shannon Index` = one_H_alpha, SI_sd = one_H_alpha_sd, `He` = two_H_alpha, `He_sd` = two_H_alpha_sd) |>
  dplyr::select(Iteration, Region, `Allelic Richness`, AR_sd, `Shannon Index`, `SI_sd`, He, `He_sd`)

# Separate the three metrics and pivot them longer
# Allelic Richness
long_ar_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Region, `Allelic Richness`, `AR_sd`) |>
  pivot_longer(
    cols = `Allelic Richness`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, AR_sd) |>
  dplyr::rename(sd = `AR_sd`)
# Shannon Index
long_SI_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Region, `Shannon Index`, `SI_sd`) |>
  pivot_longer(
    cols = `Shannon Index`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, SI_sd) |>
  dplyr::rename(sd = `SI_sd`)
# He
long_He_Hot <- diversity_combined_Hot |>
  dplyr::select(Iteration, Region, `He`, `He_sd`) |>
  pivot_longer(
    cols = `He`,
    names_to = "Metric",
    values_to = "Value"
  ) |>
  dplyr::select(Iteration, Metric, Region, Value, He_sd) |>
  dplyr::rename(sd = `He_sd`)

# Combine three metrics
long_diversity_Hot<- rbind(long_ar_Hot, long_SI_Hot, long_He_Hot)

# Calculate SE
long_diversity_Hot<- long_diversity_Hot |> mutate(se = sd/sqrt(48))

diversity_sum_Hot <- summarySE(long_diversity_Hot, measurevar="Value",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::select(Metric, Region, mean)

diversity_se_Hot <- summarySE(long_diversity_Hot, measurevar="se",
                     groupvars=c("Metric","Region"), na.rm=TRUE) |>
  dplyr::rename(SE = mean) |>
  dplyr::select(SE)

diversity_hot <- cbind(diversity_sum_Hot, diversity_se_Hot) |>
  mutate(Temp = "Hot")
```

```{r}
diversity_summary_region <- rbind(diversity_pre, diversity_amb, diversity_hot)
```

```{r}
#saveRDS(diversity_summary_region, file = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/diversity_metrics/diversity_summary_region.rds")
diversity_summary_region <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/diversity_metrics/diversity_summary_region.rds") |>
  mutate(Temp = factor(Temp, levels = c("Pre","Amb","Hot")))
```

```{r}
div_reg_plot<- diversity_summary_region |>
  ggplot(aes(x=Region, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  #scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
div_reg_plot
```

```{r}
# Save results to calculate difference in Pre diversity metrics
# Allelic Richness
mean_central_AR <- diversity_summary_region$mean[1]
mean_hybrid_AR <- diversity_summary_region$mean[4]
mean_north_AR <- diversity_summary_region$mean[7]
se_central_AR <- diversity_summary_region$SE[1]
se_hybrid_AR <- diversity_summary_region$SE[4]
se_north_AR <- diversity_summary_region$SE[7]

#AR central hybrid
mean_hybrid_AR - mean_central_AR
sqrt(se_hybrid_AR^2 + se_central_AR^2)
#AR north hybrid
mean_hybrid_AR - mean_north_AR
sqrt(se_hybrid_AR^2 + se_north_AR^2)
#AR central north
mean_north_AR - mean_central_AR
sqrt(se_central_AR^2 + se_central_AR^2)

#Heterozygosity
mean_central_HE <- diversity_summary_region$mean[1]
mean_hybrid_HE <- diversity_summary_region$mean[4]
mean_north_HE <- diversity_summary_region$mean[7]
se_central_HE <- diversity_summary_region$SE[1]
se_hybrid_HE <- diversity_summary_region$SE[4]
se_north_HE <- diversity_summary_region$SE[7]

#HE central hybrid
mean_hybrid_HE - mean_central_HE
sqrt(se_hybrid_HE^2 + se_central_HE^2)
#HE north hybrid
mean_hybrid_HE - mean_north_HE
sqrt(se_hybrid_HE^2 + se_north_HE^2)
#HE central north
mean_north_HE - mean_central_HE
sqrt(se_central_HE^2 + se_central_HE^2)

mean_central_SI <- diversity_summary_region$mean[1]
mean_hybrid_SI <- diversity_summary_region$mean[4]
mean_north_SI <- diversity_summary_region$mean[7]
se_central_SI <- diversity_summary_region$SE[1]
se_hybrid_SI <- diversity_summary_region$SE[4]
se_north_SI <- diversity_summary_region$SE[7]

#SI central hybrid
mean_hybrid_SI - mean_central_SI
sqrt(se_hybrid_SI^2 + se_central_SI^2)
#SI north hybrid
mean_hybrid_SI - mean_north_SI
sqrt(se_hybrid_SI^2 + se_north_SI^2)
#SI central north
mean_north_SI - mean_central_SI
sqrt(se_central_SI^2 + se_central_SI^2)
```

####
Heterozygosity
```{r}
#het2<- gl.report.heterozygosity(ten_larvae, plot.out = F)
het<- read.table("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/diversity_metrics/heterosis.txt", header = T) |>
  mutate(temp = factor(temp, levels = c("Pre", "Amb", "Hot"))) |>
  dplyr::filter(mod.sing == "sing") |>
  mutate(region_temp = paste(region.cross2, temp, sep = "_")) |>
  mutate(reef_temp = paste(reef.cross, temp, sep = "_")) |>
  mutate(metric = "Ho")
```

```{r}
het_box<- het |>
  ggplot(aes(x = region_temp, y = Ho, fill = region.cross2)) +
  geom_boxplot() +
  facet_wrap(~metric) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  labs(x = "Region & Temperature", y = "Heterozygosity", fill = "Region") +
  theme_Publication()+
  #scale_fill_manual(values=c("#386cb0","#7fc97f","#fdb462"))+
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
    scale_x_discrete(limits = c("Central_Pre","Central_Amb", "Central_Hot", "Hybrid_Pre", "Hybrid_Amb", "Hybrid_Hot", "North_Pre", "North_Amb", "North_Hot"))+
    scale_fill_manual(values=pal[c(1,3,5)]) +
  scale_y_continuous(limits = c(0.04, 0.12), breaks = seq(0.04, 0.12, 0.02))
het_box
```

```{r}
#stat test for variance per temp treatment
het_sum <- summarySE(het, measurevar="Ho",
                     groupvars=c("region.cross2","temp"), na.rm=TRUE)

#normally distributed?
ks.test(het$Ho, "pnorm", mean=mean(het$Ho), sd=sd(het$Ho))
qqnorm(het$Ho)
qqline(het$Ho, col = "red")

#plot
het|> ggplot() + geom_boxplot(aes(x=temp, y=Ho))
#not normally distributed

##Variance between Temp treatments
#test dispersion in all temp treatment between Pre, 27.5C and 35.5C
het_pre<- het |> dplyr::filter(temp == "Pre")
het_amb<- het |> dplyr::filter(temp == "Amb")
het_hot<- het |> dplyr::filter(temp == "Hot")
het_amb_hot<- het |> dplyr::filter(temp != "Pre")
#F test
var.test(het_pre$Ho, het_amb$Ho)
var.test(het_pre$Ho, het_hot$Ho)
var.test(het_amb$Ho, het_hot$Ho)

##Dispersion between Temp treatments and regions
het_pre_cen<- het_pre |> dplyr::filter(region.cross == "Central")
het_pre_hyb<- het_pre |> dplyr::filter(region.cross == "Hybrid")
het_pre_nor<- het_pre |> dplyr::filter(region.cross == "North")
het_amb_cen<- het_amb |> dplyr::filter(region.cross == "Central")
het_amb_hyb<- het_amb |> dplyr::filter(region.cross == "Hybrid")
het_amb_nor<- het_amb |> dplyr::filter(region.cross == "North")
het_hot_cen<- het_hot |> dplyr::filter(region.cross == "Central")
het_hot_hyb<- het_hot |> dplyr::filter(region.cross == "Hybrid")
het_hot_nor<- het_hot |> dplyr::filter(region.cross == "North")

#test dispersion between 27.5C and 35.5C in central, hybrid and north crosses
#F test
var.test(het_hot_cen$Ho, het_amb_cen$Ho)
var.test(het_hot_hyb$Ho, het_amb_hyb$Ho)
var.test(het_hot_nor$Ho, het_amb_nor$Ho)

#test dispersion between Pre and 27.5C in central, hybrid and north crosses
#F test
var.test(het_pre_cen$Ho, het_amb_cen$Ho)
var.test(het_pre_hyb$Ho, het_amb_hyb$Ho)
var.test(het_pre_nor$Ho, het_amb_nor$Ho)

##Dispersion between regions
#test dispersion of 35.5C treatment between central, hybrid and north crosses
#F test
var.test(het_hot_cen$Ho, het_hot_hyb$Ho)
var.test(het_hot_cen$Ho, het_hot_nor$Ho)
var.test(het_hot_nor$Ho, het_hot_hyb$Ho)

#F test Pre treatment comparison
var.test(het_pre_cen$Ho, het_pre_hyb$Ho)
var.test(het_pre_cen$Ho, het_pre_nor$Ho)
var.test(het_pre_nor$Ho, het_pre_hyb$Ho)
```

```{r}
# Perform Kruskal-Wallis test Ho Pre heat
#by region cross
kw_cross_pre <- kruskal.test(Ho ~ region.cross2, data = het_pre)
kw_cross_pre

# If significant, perform post hoc pairwise comparisons
if (kw_cross_pre$p.value < 0.05) {
  # Use the pairwise Wilcoxon test for post hoc analysis
  posthoc <- pairwise.wilcox.test(het_pre$Ho, het_pre$region.cross2, p.adjust.method = "none")
  print(posthoc)
} else {
  print("No significant differences, skipping post hoc analysis.")
}

# Perform Kruskal-Wallis test Ho Amb treatment
#by region cross
kw_cross_amb <- kruskal.test(Ho ~ region.cross, data = het_amb)
kw_cross_amb

# If significant, perform post hoc pairwise comparisons
if (kw_cross_amb$p.value < 0.05) {
  # Use the pairwise Wilcoxon test for post hoc analysis
  posthoc <- pairwise.wilcox.test(het_amb$Ho, het_amb$region.cross2, p.adjust.method = "none")
  print(posthoc)
} else {
  print("No significant differences, skipping post hoc analysis.")
}

# Perform Kruskal-Wallis test Ho Hot treatment
#by region cross
kw_cross_hot <- kruskal.test(Ho ~ region.cross, data = het_hot)
kw_cross_hot

# If significant, perform post hoc pairwise comparisons
if (kw_cross_hot$p.value < 0.05) {
  # Use the pairwise Wilcoxon test for post hoc analysis
  posthoc <- pairwise.wilcox.test(het_hot$Ho, het_hot$region.cross2, p.adjust.method = "none")
  print(posthoc)
} else {
  print("No significant differences, skipping post hoc analysis.")
}

# Perform Kruskal-Wallis test Ho Hot treatment
#by region cross
kw_cross_amb_hot <- kruskal.test(Ho ~ region.cross, data = het_amb_hot)
kw_cross_amb_hot

# If significant, perform post hoc pairwise comparisons
if (kw_cross_amb_hot$p.value < 0.05) {
  # Use the pairwise Wilcoxon test for post hoc analysis
  posthoc <- pairwise.wilcox.test(het_amb_hot$Ho, het_amb_hot$region.cross2, p.adjust.method = "none")
  print(posthoc)
} else {
  print("No significant differences, skipping post hoc analysis.")
}

het_amb_hot_cen<- het_amb_hot |> filter(region.cross2 == "Central")
het_amb_hot_hyb<- het_amb_hot |> filter(region.cross2 == "Hybrid")
het_amb_hot_nor<- het_amb_hot |> filter(region.cross2 == "North")

# Perform Kruskal-Wallis test Ho temp Central treatment
#by region cross
kw_cross_amb_hot_cen <- kruskal.test(Ho ~ temp, data = het_amb_hot_cen)
kw_cross_amb_hot_cen

# Perform Kruskal-Wallis test Ho temp Hybrid treatment
#by region cross
kw_cross_amb_hot_hyb <- kruskal.test(Ho ~ temp, data = het_amb_hot_hyb)
kw_cross_amb_hot_hyb

# Perform Kruskal-Wallis test Ho temp North treatment
#by region cross
kw_cross_amb_hot_nor <- kruskal.test(Ho ~ temp, data = het_amb_hot_nor)
kw_cross_amb_hot_nor
```

```{r}
library(patchwork)
a<- div_plot2 + het_box +plot_annotation(tag_levels = "A") + plot_layout(nrow = 1, guides = "collect", widths = c(5,1))
a
```

```{r}
ar<- diversity_summary |> filter(Metric == "Allelic Richness")
si<- diversity_summary |> filter(Metric == "Shannon Index")
he<- diversity_summary |> filter(Metric == "He")

ar_plot<- ar |>
  ggplot(aes(x=Reefcross, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
ar_plot

si_plot<- si |>
  ggplot(aes(x=Reefcross, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
si_plot

he_plot<- he |>
  ggplot(aes(x=Reefcross, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
he_plot

b<- ar_plot + si_plot + he_plot + het_box +plot_annotation(tag_levels = "A") + plot_layout(nrow = 2, guides = "collect")
b
```

```{r}
div_plot3<- diversity_region

ar<- diversity_region |> filter(Metric == "Allelic Richness")
si<- diversity_region |> filter(Metric == "Shannon Index")
he<- diversity_region |> filter(Metric == "He")

ar_plot<- ar |>
  ggplot(aes(x=Region, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  #scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
ar_plot

si_plot<- si |>
  ggplot(aes(x=Region, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  #scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
si_plot

he_plot<- he |>
  ggplot(aes(x=Region, y=mean))+
  geom_pointrange(aes(ymin=`mean`-`SE`, ymax=`mean`+`SE`, color=Region, group = Temp, shape = Temp), position=position_dodge(width=0.6))+
  facet_wrap(~Metric, scales = "free_y") +
  labs(x='Reef Cross', y="Diversity")+
  theme_Publication() +
  theme(legend.key.height = unit(0.3, "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,0,0,0),
        plot.margin = margin(rep(0.5,4)),
        legend.position="right",
        legend.direction = "vertical",
        axis.text.x = element_text(angle = 90),
        axis.text.y = element_text(angle = 90),
        strip.background = element_rect(fill = "white", color = "black")
        ) +
  #scale_x_discrete(limits = c("DavxDav","PalxPal","PalxDav","JewxDav","JewxPal","SwixDav", "JewxJew", "ParxPar", "SwixSwi")) +
    scale_color_manual(values=pal[c(1,3,5)])
he_plot

b<- ar_plot + si_plot + he_plot + het_box +plot_annotation(tag_levels = "A") + plot_layout(nrow = 2, guides = "collect")
b
```

```{r}
b<- ar_plot + si_plot + he_plot + het_box +plot_annotation(tag_levels = "A") + plot_layout(nrow = 1, guides = "collect")
b

b<- ar_plot + si_plot + he_plot + het_box + plot_layout(nrow = 1, guides = "collect")
b
```
