---
title: "06_A.ten_larvae_survival_correlation"
output: html_document
date: "2024-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and tools
```{r}
library(tidyverse)
library(ggpubr)
library(dartR)
library(wesanderson)
library(janitor)
library(glmnet)

set.seed(123)

#colour palette
pal <- wes_palette("Zissou1", 5, type = "discrete")

#theme for plots
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

# Function for summarising survival standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

Load data
```{r}
#survival data
survival<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/aten_larvae_survival_corr.csv")
#dart data
ten_larvae <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_larvae_filter.RDS")
#keep T0
ten_larvae_T0<- gl.keep.pop(ten_larvae, pop.list = c('T0'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
#drop models
ten_larvae_T0<- gl.drop.pop(ten_larvae_T0, pop.list = c('Mod1','Mod2','Mod3','Mod4','Mod5','Mod6'), as.pop = 'family', mono.rm = FALSE, recalc=TRUE)
```

Create objects for each family
```{r}
Dav1xDav8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav1xDav8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Dav6xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav6xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Dav8xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Dav8xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew2xJew8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew2xJew8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew2xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew2xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew3xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew3xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew6xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew6xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav10 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav10"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav13 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav13"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xJew2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xJew2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xJew3 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xJew3"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Jew8xPal1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Jew8xPal1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal2xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal2xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal2xPal5 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal2xPal5"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal4xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal4xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal4xPal2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal4xPal2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Pal5xPal4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Pal5xPal4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par2xPar5 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par2xPar5"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par4xPar2 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par4xPar2"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par5xDav6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par5xDav6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Par5xPar4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Par5xPar4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi4xDav8 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi4xDav8"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi4xSwi7 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi4xSwi7"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi6xDav1 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi6xDav1"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi6xSwi4 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi6xSwi4"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi7xDav10 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi7xDav10"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
Swi7xSwi6 <- gl.keep.pop(ten_larvae_T0, pop.list= c("Swi7xSwi6"), as.pop = "indcross", mono.rm = FALSE, recalc=TRUE)
```

Create dataframe of allele frequency for each family
```{r}
Dav1xDav8_alf <- gl.alf(Dav1xDav8) |> rownames_to_column() |> dplyr::rename(Dav1xDav8 = alf1) |> dplyr::select(Dav1xDav8)
Dav6xDav1_alf <- gl.alf(Dav6xDav1) |> rownames_to_column() |> dplyr::rename(Dav6xDav1 = alf1) |> dplyr::select(Dav6xDav1)
Dav8xDav6_alf <- gl.alf(Dav8xDav6) |> rownames_to_column() |> dplyr::rename(Dav8xDav6 = alf1) |> dplyr::select(Dav8xDav6)
Jew2xJew8_alf <- gl.alf(Jew2xJew8) |> rownames_to_column() |> dplyr::rename(Jew2xJew8 = alf1) |> dplyr::select(Jew2xJew8)
Jew2xPal1_alf <- gl.alf(Jew2xPal1) |> rownames_to_column() |> dplyr::rename(Jew2xPal1 = alf1) |> dplyr::select(Jew2xPal1)
Jew3xPal1_alf <- gl.alf(Jew3xPal1) |> rownames_to_column() |> dplyr::rename(Jew3xPal1 = alf1) |> dplyr::select(Jew3xPal1)
Jew6xPal1_alf <- gl.alf(Jew6xPal1) |> rownames_to_column() |> dplyr::rename(Jew6xPal1 = alf1) |> dplyr::select(Jew6xPal1)
Jew8xDav1_alf <- gl.alf(Jew8xDav1) |> rownames_to_column() |> dplyr::rename(Jew8xDav1 = alf1) |> dplyr::select(Jew8xDav1)
Jew8xDav10_alf <- gl.alf(Jew8xDav10) |> rownames_to_column() |> dplyr::rename(Jew8xDav10 = alf1) |> dplyr::select(Jew8xDav10)
Jew8xDav13_alf <- gl.alf(Jew8xDav13) |> rownames_to_column() |> dplyr::rename(Jew8xDav13 = alf1) |> dplyr::select(Jew8xDav13)
Jew8xDav6_alf <- gl.alf(Jew8xDav6) |> rownames_to_column() |> dplyr::rename(Jew8xDav6 = alf1) |> dplyr::select(Jew8xDav6)
Jew8xJew2_alf <- gl.alf(Jew8xJew2) |> rownames_to_column() |> dplyr::rename(Jew8xJew2 = alf1) |> dplyr::select(Jew8xJew2)
Jew8xJew3_alf <- gl.alf(Jew8xJew3) |> rownames_to_column() |> dplyr::rename(Jew8xJew3 = alf1) |> dplyr::select(Jew8xJew3)
Jew8xPal1_alf <- gl.alf(Jew8xPal1) |> rownames_to_column() |> dplyr::rename(Jew8xPal1 = alf1) |> dplyr::select(Jew8xPal1)
Pal2xDav1_alf <- gl.alf(Pal2xDav1) |> rownames_to_column() |> dplyr::rename(Pal2xDav1 = alf1) |> dplyr::select(Pal2xDav1)
Pal2xPal5_alf <- gl.alf(Pal2xPal5) |> rownames_to_column() |> dplyr::rename(Pal2xPal5 = alf1) |> dplyr::select(Pal2xPal5)
Pal4xDav1_alf <- gl.alf(Pal4xDav1) |> rownames_to_column() |> dplyr::rename(Pal4xDav1 = alf1) |> dplyr::select(Pal4xDav1)
Pal4xPal2_alf <- gl.alf(Pal4xPal2) |> rownames_to_column() |> dplyr::rename(Pal4xPal2 = alf1) |> dplyr::select(Pal4xPal2)
Pal5xPal4_alf <- gl.alf(Pal5xPal4) |> rownames_to_column() |> dplyr::rename(Pal5xPal4 = alf1) |> dplyr::select(Pal5xPal4)
Par2xPar5_alf <- gl.alf(Par2xPar5) |> rownames_to_column() |> dplyr::rename(Par2xPar5 = alf1) |> dplyr::select(Par2xPar5)
Par4xPar2_alf <- gl.alf(Par4xPar2) |> rownames_to_column() |> dplyr::rename(Par4xPar2 = alf1) |> dplyr::select(Par4xPar2)
Par5xDav6_alf <- gl.alf(Par5xDav6) |> rownames_to_column() |> dplyr::rename(Par5xDav6 = alf1) |> dplyr::select(Par5xDav6)
Par5xPar4_alf <- gl.alf(Par5xPar4) |> rownames_to_column() |> dplyr::rename(Par5xPar4 = alf1) |> dplyr::select(Par5xPar4)
Swi4xDav8_alf <- gl.alf(Swi4xDav8) |> rownames_to_column() |> dplyr::rename(Swi4xDav8 = alf1) |> dplyr::select(Swi4xDav8)
Swi4xSwi7_alf <- gl.alf(Swi4xSwi7) |> rownames_to_column() |> dplyr::rename(Swi4xSwi7 = alf1) |> dplyr::select(Swi4xSwi7)
Swi6xDav1_alf <- gl.alf(Swi6xDav1) |> rownames_to_column() |> dplyr::rename(Swi6xDav1 = alf1) |> dplyr::select(Swi6xDav1)
Swi6xSwi4_alf <- gl.alf(Swi6xSwi4) |> rownames_to_column() |> dplyr::rename(Swi6xSwi4 = alf1) |> dplyr::select(Swi6xSwi4)
Swi7xDav10_alf <- gl.alf(Swi7xDav10) |> rownames_to_column() |> dplyr::rename(Swi7xDav10 = alf1) |> dplyr::select(Swi7xDav10)
Swi7xSwi6_alf <- gl.alf(Swi7xSwi6) |> rownames_to_column() |> dplyr::rename(Swi7xSwi6 = alf1) |> dplyr::select(Swi7xSwi6)
```

extract SNP names
```{r}
SNPs <- gl.alf(Dav1xDav8) |> rownames_to_column() |> dplyr::rename(SNPs = rowname) |> dplyr::select(SNPs)
```

Bind all allele frequencies together
```{r}
alf<- cbind(
SNPs,
Dav1xDav8_alf$Dav1xDav8, 
Dav6xDav1_alf$Dav6xDav1, 
Dav8xDav6_alf$Dav8xDav6, 
Jew2xJew8_alf$Jew2xJew8, 
Jew2xPal1_alf$Jew2xPal1, 
Jew3xPal1_alf$Jew3xPal1, 
Jew6xPal1_alf$Jew6xPal1, 
Jew8xDav1_alf$Jew8xDav1, 
Jew8xDav10_alf$Jew8xDav10, 
Jew8xDav13_alf$Jew8xDav13, 
Jew8xDav6_alf$Jew8xDav6, 
Jew8xJew2_alf$Jew8xJew2, 
Jew8xJew3_alf$Jew8xJew3, 
Jew8xPal1_alf$Jew8xPal1, 
Pal2xDav1_alf$Pal2xDav1, 
Pal2xPal5_alf$Pal2xPal5, 
Pal4xDav1_alf$Pal4xDav1, 
Pal4xPal2_alf$Pal4xPal2, 
Pal5xPal4_alf$Pal5xPal4, 
Par2xPar5_alf$Par2xPar5, 
Par4xPar2_alf$Par4xPar2, 
Par5xDav6_alf$Par5xDav6, 
Par5xPar4_alf$Par5xPar4, 
Swi4xDav8_alf$Swi4xDav8, 
Swi4xSwi7_alf$Swi4xSwi7, 
Swi6xDav1_alf$Swi6xDav1, 
Swi6xSwi4_alf$Swi6xSwi4, 
Swi7xDav10_alf$Swi7xDav10, 
Swi7xSwi6_alf$Swi7xSwi6
)
```

Transpose and create rownames
```{r}
alf<- alf |> 
  t() |>
  row_to_names(row_number = 1)
```

Change class of matrix
```{r}
class(alf) <- "numeric"
alf2<- as.data.frame(alf)
```

Correlation between SNP alf and Larval survival
```{r}
#run correlation analysis
cor1<-psych::corr.test(alf2[c(1:13057)], survival$Survival, method="pearson", adjust = "BH")
ci<- cor1$ci2
padj<- cor1$p.adj
```

Make datframes of highly significant SNPs
```{r}
#filter highly predictive
pred_SNPs<- ci |> dplyr::filter(r > 0.5 | r < -0.5)
highly_pred_SNPs_0.1<- ci |> dplyr::filter(pa < 0.1)
highly_pred_SNPs<- ci |> dplyr::filter(pa < 0.05)

#extract SNP names (ci for some reason butchers the names)
pearson_snps_0.5<- highly_pred_SNPs |> rownames_to_column(var = "SNP")
pearson_snps_0.5<- pearson_snps_0.5$SNP
pearson_snps_0.5<- gsub("-NA", "", pearson_snps_0.5)
SNP_list<- SNPs$SNPs |> as.vector()
pattern <- paste(pearson_snps_0.5, collapse = "|")
pearson_snps_0.5<- SNP_list[grep(pattern, SNP_list)]

pearson_snps_1.0<- highly_pred_SNPs_0.1 |> rownames_to_column(var = "SNP")
pearson_snps_1.0<- pearson_snps_1.0$SNP
pearson_snps_1.0<- gsub("-NA", "", pearson_snps_1.0)
SNP_list<- SNPs$SNPs |> as.vector()
pattern <- paste(pearson_snps_1.0, collapse = "|")
pearson_snps_1.0<- SNP_list[grep(pattern, SNP_list)]
```

Plot correlations
```{r}
#plot positive correlations
SNP_15476258<- alf2 |> dplyr::select(`15476258-41-A/G`)
SNP_15476258<- cbind(SNP_15476258, survival)
plot15476258<- ggplot(SNP_15476258, aes(x = `15476258-41-A/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='glm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15476258

SNP_15477968<- alf2 |> dplyr::select(`15477968-19-C/A`)
SNP_15477968<- cbind(SNP_15477968, survival)
plot15477968<-ggplot(SNP_15477968, aes(x = `15477968-19-C/A`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15477968

SNP_73055070<- alf2 |> dplyr::select(`73055070-13-A/G`)
SNP_73055070<- cbind(SNP_73055070, survival)
plot73055070<-ggplot(SNP_73055070, aes(x = `73055070-13-A/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot73055070

SNP_15457915<- alf2 |> dplyr::select(`15457915-49-G/C`)
SNP_15457915<- cbind(SNP_15457915, survival)
plot15457915<-ggplot(SNP_15457915, aes(x = `15457915-49-G/C`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot15457915

#plot negative correlations
SNP_9498501<- alf2 |> dplyr::select(`9498501-47-C/T`)
SNP_9498501<- cbind(SNP_9498501, survival)
plot9498501<-ggplot(SNP_9498501, aes(x = `9498501-47-C/T`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot9498501

SNP_11466799<- alf2 |> dplyr::select(`11466799-26-T/G`)
SNP_11466799<- cbind(SNP_11466799, survival)
plot11466799<-ggplot(SNP_11466799, aes(x = `11466799-26-T/G`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot11466799

SNP_11461901<- alf2 |> dplyr::select(`11461901-9-T/C`)
SNP_11461901<- cbind(SNP_11461901, survival)
plot11461901<- ggplot(SNP_11461901, aes(x = `11461901-9-T/C`, y = Survival)) +
  geom_point(aes(colour = Region), size = 2) +
  geom_smooth(method='lm', formula= y~x, se = FALSE, colour = "black")+
  scale_color_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  labs(x="Allele Frequency")+
  theme(legend.direction = "vertical")
plot11461901

#combine plots
library(patchwork)
plot15476258 + plot15477968 + plot73055070 + plot9498501 + plot11466799 + plot11461901 + plot_layout(ncol = 3, guides = 'collect') + plot_annotation(tag_levels = 'A')
```

Find and plot significant SNPs to see differences in AF between region crosses
```{r}
#filter out the significant SNPs
sig_als<- padj |> as.data.frame() |> dplyr::filter(V1 < 0.05) |> rownames_to_column(var = "SNPs")
sig_als <- sig_als$SNPs |> as.vector()
sig_als2<- alf2 |> dplyr::select(all_of(sig_als))

#bind survival and allele frequency together
sig_als3<- cbind(survival, sig_als2)
rownames(sig_als3) <- NULL

#Create dataframe of sig NSP AF per region
summary_AF <- sig_als3 |>
  group_by(Region) |>
  dplyr::summarise(across(contains("/"), mean, na.rm = TRUE)) |>
  pivot_longer(cols = contains("/"), 
               names_to = "SNP", 
               values_to = "allele_frequency")

# Plot violin plot of allele frequencies by region
vio_af_plot<- ggplot(summary_AF, aes(x = Region, y = allele_frequency, colour = Region)) +
  geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") +
  scale_color_manual(values=pal[c(1,3,5)]) +
  scale_fill_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
vio_af_plot

#change AF from 0-1 to 0-0.5
summary_AF2<- summary_AF |>
  mutate(alf2 = ifelse(allele_frequency > 0.5, 1 - allele_frequency, allele_frequency))

#(Looks better the other way eg in relation to the reference SNP)
vio_af_plot<- ggplot(summary_AF2, aes(x = Region, y = alf2, colour = Region)) +
  geom_violin(aes(x = Region, y = alf2, fill = Region), alpha = 0.3, adjust = 10) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1) +  # Points with slight jitter for clarity
  labs(x = "Region", y = "Allele Frequency") +
  scale_color_manual(values=pal[c(1,3,5)]) +
  scale_fill_manual(values=pal[c(1,3,5)]) +
  theme_Publication()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
vio_af_plot
```

Statistics to see if AF is different between regions (KS test)
```{r}
# Separate the data by region
central_af <- summary_AF %>% filter(Region == "Central") %>% pull(allele_frequency)
hybrid_af <- summary_AF %>% filter(Region == "Hybrid") %>% pull(allele_frequency)
north_af <- summary_AF %>% filter(Region == "North") %>% pull(allele_frequency)

# Perform KS test between Central and Hybrid regions
ks_test_central_hybrid <- ks.test(central_af, hybrid_af)
ks_test_central_hybrid

# Perform KS test between Central and North regions
ks_test_central_north <- ks.test(central_af, north_af)
ks_test_central_north

# Perform KS test between Hybrid and North regions
ks_test_hybrid_north <- ks.test(hybrid_af, north_af)
ks_test_hybrid_north

#means of AF per region
AF_means <- summarySE(summary_AF, measurevar="allele_frequency",
                     groupvars=c("Region"), na.rm=TRUE)
```

Line plot of SNPs
```{r}
# Calculate the standard deviation of allele frequencies by SNP across regions
variation_summary <- summary_AF %>% dplyr::filter(Region != "Hybrid") |>
  group_by(SNP) %>%
  dplyr::summarise(var = sd(allele_frequency)) %>%
  dplyr::arrange(desc(var))

# Filter for the top SNPs with highest variation
top_SNPs <- variation_summary %>%
  filter(var > quantile(var, 0.75)) %>%  # Keep SNPs with variation in the top 25%
  pull(SNP)

# Filter the original data for top SNPs
highlighted_data <- summary_AF %>% 
  filter(SNP %in% top_SNPs)

# Plot with highlighted SNPs
lin_af_plot<-ggplot(highlighted_data, aes(x = Region, y = allele_frequency, colour = Region, group = SNP)) + 
  #geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1.5) +  # Points with slight jitter for clarity
  geom_line(aes(group = SNP), colour = "gray", size = 0.8, alpha = 0.5) +  # Lines connecting points for same SNP
  labs(x = "Region", y = "Allele Frequency") + 
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lin_af_plot

#change AF from 0-1 to 0-0.5
highlighted_data2<- highlighted_data |>
  mutate(alf2 = ifelse(allele_frequency > 0.5, 1 - allele_frequency, allele_frequency))

# (looks better the ref based way)
lin_af_plot<-ggplot(highlighted_data2, aes(x = Region, y = alf2, colour = Region, group = SNP)) + 
  #geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1.5) +  # Points with slight jitter for clarity
  geom_line(aes(group = SNP), colour = "gray", size = 0.8, alpha = 0.5) +  # Lines connecting points for same SNP
  labs(x = "Region", y = "Allele Frequency") + 
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lin_af_plot
```

```{r}
# Filter for the top SNPs with highest variation
cen_SNPs <- summary_AF %>%
  #dplyr::filter(Region == "Hybrid") |>
  dplyr::filter(Region == "Central" & (allele_frequency < 0.2 | allele_frequency > 0.85)) %>%
  pull(SNP)
hyb_SNPs <- summary_AF %>%
  #dplyr::filter(Region == "Hybrid") |>
  dplyr::filter(Region == "Hybrid" & allele_frequency > 0.35 & allele_frequency < 0.65) %>%
  pull(SNP)

top_SNPs <- intersect(cen_SNPs, hyb_SNPs)

# Filter the original data for top SNPs
highlighted_data <- summary_AF %>% 
  filter(SNP %in% top_SNPs)

# Plot with highlighted SNPs
lin_af_plot<-ggplot(highlighted_data, aes(x = Region, y = allele_frequency, colour = Region, group = SNP)) + 
  #geom_violin(aes(x = Region, y = allele_frequency, fill = Region), alpha = 0.3) +  # Violin plot for density
  geom_point(position = position_jitter(width = 0.1, height = 0), size = 1.5) +  # Points with slight jitter for clarity
  geom_line(aes(group = SNP), colour = "gray", size = 0.8, alpha = 0.5) +  # Lines connecting points for same SNP
  labs(x = "Region", y = "Allele Frequency") + 
  scale_color_manual(values = pal[c(1, 3, 5)]) + 
  scale_fill_manual(values = pal[c(1, 3, 5)]) + 
  theme_Publication() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lin_af_plot
```

Manhatten plot
```{r}
# Chromosome length info
lengths <- read_tsv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ragtag.scaffold_lengths.txt", col_names = c("chr","length"),show_col_types = FALSE)

#load chromosome offsets
offsets <- lengths %>% arrange(dplyr::desc(length)) %>% 
  dplyr::mutate(offset=cumsum(length)-length) %>% 
  dplyr::mutate(scaffold_num = row_number())

axis_chr<- offsets %>% 
  mutate(centre=offset+length/2) %>% 
  mutate(chr_id=ifelse(grepl(chr,pattern="chr"),chr,"Unplaced")) %>% 
  mutate(chr_id=str_remove(chr_id,"_RagTag"),chr_id=str_remove(chr_id,"chr")) %>% 
  group_by(chr_id) %>% 
  dplyr::summarise(centre = mean(centre))
```

```{r}
#create -log10 column
ci2<- ci |> mutate(`-log10(p)` = -log10(p)) |> mutate(`-log10(p)` = coalesce(`-log10(p)`, 0.01))
#write.table(ci2, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ci.txt", quote=F, row.names=F, sep = "\t")
```

### Conversion to pseudo-chromosome coordinates

All outputs were converted to RagTag pseudo-chromosome coordinates.  The process for is illustrated below

In this case the input example has four columns. The first are chromosome and position.  Then there are two more data columns.  You can have as many data columns as you like. 

NOTE: remove column names from dataframe before this step

```{bash}
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/manhattan_plotting

#python ./translate_coords.py --keep ci.txt ragtag.scaffold.agp > ci.ragtag.txt
```

```{r}
# load manhatten plot data
mh_data_corr <- read_table("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ci.ragtag.txt",col_names = c("chr","pos","SNP","lower","r","upper","p","pa","-log10(p)","chr_orig","pos_orig")) %>% 
  left_join(offsets) %>%
  mutate(abs_pos = pos + offset) %>%
  mutate(chr_type=ifelse(scaffold_num %% 2 == 0,"dark", "light")) %>% 
  mutate(chr_type=ifelse(grepl(chr,pattern="chr"),chr_type,"unplaced"))
```

```{r}
# colour palette
ccols <- hcl.colors(n=3,palette = "Tropic")

#plot manhatten plot
mh_plot_corr<- mh_data_corr %>% 
  ggplot() + 
  geom_point(aes(x=abs_pos/1e+6,y=`-log10(p)`,color=chr_type),size=1,alpha=.6) + 
  scale_x_continuous(label = axis_chr$chr_id, breaks = axis_chr$centre/1e6) +
  geom_hline(yintercept = 3.825559, size = 0.6) +
  #geom_hline(yintercept = 5.1, size = 0.6) +
  xlab("Chromosome") +
  ylab("-log10(p)") + 
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  guides(color="none") +
  #scale_color_manual(values = c("light"=ccols[1],"dark"="grey","unplaced"=ccols[3]))
  scale_color_manual(values = c("light"="#e8998d","dark"="#6c9a8b","unplaced"="darkgrey"))
mh_plot_corr
```

Find genes that SNPs are in
```{r}
#filter highly predictive
sig_SNPs<- mh_data_corr |> dplyr::filter(pa < 0.1)
highly_sig_SNPs<- mh_data_corr |> dplyr::filter(pa < 0.05)

#select only the SNP column
sig_SNPs<- sig_SNPs |> dplyr::select(SNP)

#remove header
names(sig_SNPs)= NULL

#write out list
#write.table(sig_SNPs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_SNPs.txt", row.names = FALSE, quote = F)
```

```{bash}
#create vcf with only significant SNPs
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data
#vcftools --vcf aten_larvae_T0.vcf --snps sig_SNPs.txt --recode --out filtered.vcf
```

```{bash}
#extract the info from the filtered vcf file
#change directory
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data
#save vcf as object
#VCF=/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/filtered.vcf
#extract first 3 columns
#grep -v "^#" $VCF | cut -f1-3 | awk '{print $0"\t"NR}' > larv_SNPs.txt
```

```{bash}
#convert filtered vcf file to bed files
#change directory
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data
#with bedops, convert vcf to bed file
#vcf2bed --snvs < filtered.vcf > filtered.bed
#save bed file as object
#BED=/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/filtered.bed
#extract first 4 columns
#grep -v "^#" $BED | cut -f1-4 | awk '{print $0"\t"NR}' > sig_SNPs.bed
#with bedops, convert gtf file to bed file of genes
#gtf2bed < aten.gtf > genes.bed
```

Flanking
```{bash}
#extract gene names that SNPs are within 10kb of
#Use BEDTools flank command to extend the SNP regions by 10 kb on both sides
#bedtools flank -i sig_SNPs.bed -g genes.bed -b 10000 > snps_10k.bed
#once you have both the SNP BED file and the gene BED file, you can use bedtools intersect to find overlaps
#bedtools intersect -a snps_10k.bed -b genes.bed > snps_flanking.bed
#you can extract the gene names as well into the bed file
#bedtools intersect -a snps_10k.bed -b genes.bed -wo > snps_genes_flanking.bed
#extract SNPs and gene names
#cut -f 4,8,9 snps_genes_flanking.bed > snps_flanking_gene_names.txt
#extract only gene names
#cut -f 3 snps_flanking_gene_names.txt > flanking_gene_names.txt
#keep only unique gene names
#sort flanking_gene_names.txt | uniq > flanking_gene_names_unique.txt
#extract protein fastas of each gene
#seqkit grep -f flanking_gene_names_unique.txt aten_0.11.maker_post_001.proteins.fasta -o sig_flanking_gene_seqs.fasta
```

Use flanking genes and see which
```{r}
#load tsv
annotation<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/annotation_table.csv")
#load significant SNP flanking genes list
sig<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/flanking_gene_names_unique.csv", header = F)
#filter only significant genes
sig_ann <- annotation %>% filter(aten_id2 %in% sig$V1)
#which ones aren't in there?
missing_ann <- sig %>% filter(!V1 %in% sig_ann$aten_id)
```

KOG
```{r}
library(KOGMWU)
library(tidyverse)
#load gene2kog data
gene2kog_aten<- read_tsv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/Atenuis_gene2kogClass.tab", col_names = F) |> na.omit() 
gene2kog_aten<- gene2kog_aten |> mutate(
  V1 = factor(X1),
  V2 = factor(X2)
) |>
  dplyr::select(V1, V2) |> as.data.frame()
```

```{r}
#significant SNPs R and p values
sig_SNPs1<- mh_data_corr |> dplyr::select(SNP, r, p)

#load gene names of flanking genes to SNPs
snps_flanking_gene_names_m1<- read.table("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/snps_flanking_gene_names.txt", header = F) |>
  dplyr::rename(SNP = V1, gene = V3)

#find gene names for sig SNPs
sig_SNPs<- sig_SNPs1 |> left_join(snps_flanking_gene_names_m1, by = "SNP") |>
  dplyr::select(gene, r) |>
  na.omit() |>
  distinct(gene, .keep_all=TRUE) |>
  as.data.frame() |>
  mutate(gene = as.factor(gene))
```

```{r}
# Analyzing D1 coral response to heat stress:
KOG=kog.mwu(sig_SNPs,gene2kog_aten)
KOG
#write.csv(KOG, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/KOG_delta_ranks.csv")
```

Plot KOG terms
```{r}
KOG_plot2<- ggplot(data = KOG, aes(x = delta.rank, y = reorder(term, delta.rank), color = padj, size = nseqs)) +
  geom_point() +
  scale_color_gradient(low = "red", high = "blue") +
  labs(x = "Delta Rank", y = "KOG Term", color = "Adjusted P-value", size = "N Seqs") +
  theme_minimal()
KOG_plot2

KOG_plot1<- ggplot(data = KOG, aes(x = reorder(term, nseqs), y = nseqs, fill = delta.rank)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient(low = "red", high = "blue") +
  facet_wrap(~ cut(padj, breaks = c(0, 0.05, 0.1, 1), labels = c("p <= 0.05", "0.05 < p <= 0.1", "p > 0.1"))) +
  labs(x = "KOG Term", y = "Number of Sequences", fill = "Delta rank") +
  theme_minimal()
KOG_plot1

KOG_plot<- KOG_plot1 + KOG_plot2
```

GO enrichment analysis
```{r}
#read in sig genes
#find gene names for sig SNPs
sig_SNPs_GO<- sig_SNPs1 |> left_join(snps_flanking_gene_names_m1, by = "SNP") |>
  dplyr::select(gene, r, p) |>
  distinct(gene, .keep_all=TRUE) |>
  as.data.frame() |>
  mutate(gene = as.factor(gene)) |>
  na.omit()

sig_SNPs_GO$p[is.na(sig_SNPs_GO$p)] <- 1

#extract gene names
genes<- gene2kog_aten |> dplyr::select(V1) |> dplyr::rename(gene=V1)
#join all gene names to significant genes
sig_genes2<- genes |> left_join(sig_SNPs_GO, by = "gene")
#add .m1 to gene names to correspond with other datasets
sig_genes2$gene <- paste0(sig_genes2$gene, ".m1")
#turn non-sig genes NAs to 0s
sig_genes2$r[is.na(sig_genes2$r)] <- 0
#select relevant columns
sig_genes3<- sig_genes2 |> dplyr::select(gene, r)
#write dataframe
#write.csv(sig_genes3, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_genes.csv")

#genes of only the significant SNPs
sig_genes4<- sig_genes2 |> dplyr::filter(p < 0.000149431) |> dplyr::select(gene, r)
#write dataframe
write.csv(sig_genes4, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_genes_only.csv")
```

```{r}
#All genes
# First, press command-D on mac or ctrl-shift-H in Rstudio and navigate to the directory containing scripts and input files. Then edit, mark and execute the following bits of code, one after another.
setwd("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data")
getwd()
# Edit these to match your data file names: 
# two columns of comma-separated values: gene id, continuous measure of significance. To perform standard GO enrichment analysis based on Fisher's exact test, use binary measure (0 or 1, i.e., either sgnificant or not).
input="all_genes.csv" #only significant genes
goAnnotations= "atenuis_gene2go.tab" # two-column, tab-delimited, one line per gene, multiple GO terms separated by semicolon. If you have multiple lines per gene, use nrify_GOtable.pl prior to running this script.
goDatabase="go.obo" # download from http://www.geneontology.org/GO.downloads.ontology.shtml
goDivision="MF" # either MF, or BP, or CC ### BP = Biological process; CC = Cellular component; MF = Molecular function
source("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/gomwu.functions.R")

# ------------- Calculating stats
# It might take a few minutes for MF and BP. Do not rerun it if you just want to replot the data with different cutoffs, go straight to gomwuPlot. If you change any of the numeric values below, delete the files that were generated in previos runs first.
 
gomwuStats(input, goDatabase, goAnnotations, goDivision,
	perlPath="perl", # replace with full path to perl executable if it is not in your system's PATH already
	largest=0.5,  # a GO category will not be considered if it contains more than this fraction of the total number of genes
	smallest=2,   # a GO category should contain at least this many genes to be considered
	clusterCutHeight=0.25, # threshold for merging similar (gene-sharing) terms. See README for details.
#	Alternative="g" # by default the MWU test is two-tailed; specify "g" or "l" of you want to test for "greater" or "less" instead. 
#	Module=TRUE,Alternative="g" # un-remark this if you are analyzing a SIGNED WGCNA module (values: 0 for not in module genes, kME for in-module genes). In the call to gomwuPlot below, specify absValue=0.001 (count number of "good genes" that fall into the module)
#	Module=TRUE # un-remark this if you are analyzing an UNSIGNED WGCNA module 
)
# do not continue if the printout shows that no GO terms pass 10% FDR.

# ----------- Plotting results
quartz()
results=gomwuPlot(input,goAnnotations,goDivision,
 	absValue=-log(0.001,10),  # genes with the measure value exceeding this will be counted as "good genes". This setting is for signed log-pvalues. Specify absValue=0.001 if you are doing Fisher's exact test for standard GO enrichment or analyzing a WGCNA module (all non-zero genes = "good genes").
 #	absValue=1, # un-remark this if you are using log2-fold changes
 	level1=0.1, # FDR threshold for plotting. Specify level1=1 to plot all GO categories containing genes exceeding the absValue.
 	level2=0.05, # FDR cutoff to print in regular (not italic) font.
 	level3=0.01, # FDR cutoff to print in large bold font.
 	txtsize=1.2,    # decrease to fit more on one page, or increase (after rescaling the plot so the tree fits the text) for better "word cloud" effect
 	treeHeight=0.5, # height of the hierarchical clustering tree
 #	colors=c("dodgerblue2","firebrick1","skyblue2","lightcoral") # these are default colors, un-remar and change if needed
 )
 # manually rescale the plot so the tree matches the text 
 # if there are too many categories displayed, try make it more stringent with level1=0.05,level2=0.01,level3=0.001.  
 
 # text representation of results, with actual adjusted p-values
 res<-results[[1]]
write.csv(res,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/GO_Aten.csv")
# ------- extracting representative GOs

# this module chooses GO terms that best represent *independent* groups of significant GO terms

pcut=1e-2 # adjusted pvalue cutoff for representative GO
hcut=0.9 # height at which cut the GO terms tree to get "independent groups". 

# plotting the GO tree with the cut level (un-remark the next two lines to plot)
quartz()
 plot(results[[2]],cex=0.6)
 abline(h=hcut,col="red")

# cutting
ct=cutree(results[[2]],h=hcut)
annots=c();ci=1
for (ci in unique(ct)) {
  message(ci)
	rn=names(ct)[ct==ci]
	obs=grep("obsolete",rn)
	if(length(obs)>0) { rn=rn[-obs] }
	if (length(rn)==0) {next}
	rr=results[[1]][rn,]
	bestrr=rr[which(rr$pval==min(rr$pval)),]
	best=1
	if(nrow(bestrr)>1) {
		nns=sub(" .+","",row.names(bestrr))
		fr=c()
		for (i in 1:length(nns)) { fr=c(fr,eval(parse(text=nns[i]))) }
		best=which(fr==max(fr))
	}
	if (bestrr$pval[best]<=pcut) { annots=c(annots,sub("\\d+\\/\\d+ ","",row.names(bestrr)[best]))}
}

mwus=read.table(paste("MWU",goDivision,input,sep="_"),header=T)
bestGOs=mwus[mwus$name %in% annots,]
bestGOs

#write.csv(mwus,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/GO_Aten_mwus.csv")
#write.csv(bestGOs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/GO_Aten_bestGOs.csv")
```

```{r}
#sig genes
# First, press command-D on mac or ctrl-shift-H in Rstudio and navigate to the directory containing scripts and input files. Then edit, mark and execute the following bits of code, one after another.
setwd("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/sig_genes_only")
getwd()
# Edit these to match your data file names: 
# two columns of comma-separated values: gene id, continuous measure of significance. To perform standard GO enrichment analysis based on Fisher's exact test, use binary measure (0 or 1, i.e., either sgnificant or not).
input="sig_genes_only.csv" #only significant genes
goAnnotations= "atenuis_gene2go.tab" # two-column, tab-delimited, one line per gene, multiple GO terms separated by semicolon. If you have multiple lines per gene, use nrify_GOtable.pl prior to running this script.
goDatabase="go.obo" # download from http://www.geneontology.org/GO.downloads.ontology.shtml
goDivision="MF" # either MF, or BP, or CC ### BP = Biological process; CC = Cellular component; MF = Molecular function
source("/Users/alexmacadam/Dropbox/PhD/Chapter4_RNA/Data/Enrichment_analysis/Larvae/gomwu.functions.R")

# ------------- Calculating stats
# It might take a few minutes for MF and BP. Do not rerun it if you just want to replot the data with different cutoffs, go straight to gomwuPlot. If you change any of the numeric values below, delete the files that were generated in previos runs first.
 
gomwuStats(input, goDatabase, goAnnotations, goDivision,
	perlPath="perl", # replace with full path to perl executable if it is not in your system's PATH already
	largest=0.5,  # a GO category will not be considered if it contains more than this fraction of the total number of genes
	smallest=2,   # a GO category should contain at least this many genes to be considered
	clusterCutHeight=0.25, # threshold for merging similar (gene-sharing) terms. See README for details.
#	Alternative="g" # by default the MWU test is two-tailed; specify "g" or "l" of you want to test for "greater" or "less" instead. 
#	Module=TRUE,Alternative="g" # un-remark this if you are analyzing a SIGNED WGCNA module (values: 0 for not in module genes, kME for in-module genes). In the call to gomwuPlot below, specify absValue=0.001 (count number of "good genes" that fall into the module)
#	Module=TRUE # un-remark this if you are analyzing an UNSIGNED WGCNA module 
)
# do not continue if the printout shows that no GO terms pass 10% FDR.

# ----------- Plotting results
quartz()
results=gomwuPlot(input,goAnnotations,goDivision,
 	absValue=-log(0.001,10),  # genes with the measure value exceeding this will be counted as "good genes". This setting is for signed log-pvalues. Specify absValue=0.001 if you are doing Fisher's exact test for standard GO enrichment or analyzing a WGCNA module (all non-zero genes = "good genes").
 #	absValue=1, # un-remark this if you are using log2-fold changes
 	level1=0.1, # FDR threshold for plotting. Specify level1=1 to plot all GO categories containing genes exceeding the absValue.
 	level2=0.05, # FDR cutoff to print in regular (not italic) font.
 	level3=0.01, # FDR cutoff to print in large bold font.
 	txtsize=1.2,    # decrease to fit more on one page, or increase (after rescaling the plot so the tree fits the text) for better "word cloud" effect
 	treeHeight=0.5, # height of the hierarchical clustering tree
 #	colors=c("dodgerblue2","firebrick1","skyblue2","lightcoral") # these are default colors, un-remar and change if needed
 )
 # manually rescale the plot so the tree matches the text 
 # if there are too many categories displayed, try make it more stringent with level1=0.05,level2=0.01,level3=0.001.  
 
 # text representation of results, with actual adjusted p-values
 res<-results[[1]]
write.csv(res,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/all_genes/GO_Aten.csv")
# ------- extracting representative GOs

# this module chooses GO terms that best represent *independent* groups of significant GO terms

pcut=1e-2 # adjusted pvalue cutoff for representative GO
hcut=0.9 # height at which cut the GO terms tree to get "independent groups". 

# plotting the GO tree with the cut level (un-remark the next two lines to plot)
quartz()
 plot(results[[2]],cex=0.6)
 abline(h=hcut,col="red")

# cutting
ct=cutree(results[[2]],h=hcut)
annots=c();ci=1
for (ci in unique(ct)) {
  message(ci)
	rn=names(ct)[ct==ci]
	obs=grep("obsolete",rn)
	if(length(obs)>0) { rn=rn[-obs] }
	if (length(rn)==0) {next}
	rr=results[[1]][rn,]
	bestrr=rr[which(rr$pval==min(rr$pval)),]
	best=1
	if(nrow(bestrr)>1) {
		nns=sub(" .+","",row.names(bestrr))
		fr=c()
		for (i in 1:length(nns)) { fr=c(fr,eval(parse(text=nns[i]))) }
		best=which(fr==max(fr))
	}
	if (bestrr$pval[best]<=pcut) { annots=c(annots,sub("\\d+\\/\\d+ ","",row.names(bestrr)[best]))}
}

mwus=read.table(paste("MWU",goDivision,input,sep="_"),header=T)
bestGOs=mwus[mwus$name %in% annots,]
bestGOs

#write.csv(mwus,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_mwus.csv")
#write.csv(bestGOs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_bestGOs.csv")
```

########
Genotype of sig SNPs in parents vs survival deviance
```{r}
#Extract only significant SNPs
#0.05
sig_list_05<- sig_SNPs1 |>
  dplyr::filter(p < 0.00014)

sig_list_05_pos<- sig_list_05 |>
  dplyr::filter(r > 0)

sig_list_05_neg<- sig_list_05 |>
  dplyr::filter(r < 0)

#0.10
sig_list_10<- sig_SNPs1 |>
  dplyr::filter(p < 0.00337)

sig_list_10_pos<- sig_list_10 |>
  dplyr::filter(r > 0)

sig_list_10_neg<- sig_list_10 |>
  dplyr::filter(r < 0)

#write.csv(sig_list_10_pos, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_10_pos.csv")
#write.csv(sig_list_10_neg, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_10_neg.csv")
#write.csv(sig_list_05_pos, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_05_pos.csv")
#write.csv(sig_list_05_neg, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_05_neg.csv")
```

```{r}
sig_list_10_pos<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_10_pos.csv")

sig_list_10_neg<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/sig_list_10_neg.csv")
```

```{r}
#load adult data
ten_adult <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_adult_filter.RDS")

# Filter out only significant SNPs

#pos 0.10
ten_adult2_pos<- gl.keep.loc(ten_adult, loc.list=c('15470345-24-C/A','9368199-30-A/T','15468417-51-A/G','15464700-55-T/G','11472264-23-C/T','17483027-9-C/A','15457567-60-T/C','26733384-64-C/A','15474679-59-T/C','15449290-60-C/T','15461869-30-G/A','15457128-50-C/A','54738551-14-A/G','15470352-48-G/A','15475129-65-C/T','15475339-31-A/C','15463894-24-T/A','15483706-19-T/C','15454357-53-C/A','15476339-21-G/A','100653226-43-G/C','68347976-16-G/C','15480032-42-A/C','83600858-7-G/A','15478027-42-A/T','17483458-13-G/A','26735296-20-T/C','17481310-48-G/T','15470469-27-T/G','15471803-65-G/T','15464910-62-C/T','15476935-65-T/C','11473040-26-G/A','11467037-7-C/T','15476258-41-A/G','17481196-9-A/G','15472350-42-C/G','15471965-54-T/A','15461829-6-C/A','15464995-37-G/C','9487429-17-C/A','15472379-23-T/A','100655061-38-A/G','15478659-19-T/C','68346615-44-G/A','26748631-8-G/A','26729068-7-G/T','17481013-65-T/A','68343947-59-G/A','17482308-32-C/G','15475106-7-G/A','15476264-13-T/G','50827897-5-C/T','26733981-26-T/C','15477895-10-A/G','15465028-11-T/C','17483190-11-T/C','15471002-32-A/G','15476531-7-C/G','15474775-23-T/G','11495556-22-A/G','11478381-21-G/T','9447179-15-A/C','15479072-67-T/C','15457063-35-A/G','11482646-56-G/A','26729415-57-T/C','15464279-20-T/C','11470459-47-T/A','11482155-24-T/C','68348080-56-A/T','11496266-19-A/G','53148777-45-G/C','53148461-56-A/T','73054819-5-A/T','15469281-61-T/A','26740334-15-T/G','15477968-19-C/A','26735144-47-T/G','15477742-34-G/T','15471547-20-A/T','50827841-44-T/A','15467146-65-A/T','15471194-67-A/T','15479271-57-C/T','26736398-67-T/G','15479929-12-C/T','11472888-8-G/C','9489413-39-A/G','15477240-6-A/T','15466554-20-G/A','17481752-54-C/A','15482284-59-A/G','26731142-21-G/A','17481982-55-G/A','11490081-64-G/C','11464175-46-A/T','73054858-10-A/C','11472783-25-A/G','54739312-40-G/A','15475112-29-C/G','15476287-33-A/T','26732165-24-C/G','26738816-11-A/G','73054451-28-G/T','15461684-66-C/T','26741700-67-G/C','11460619-53-G/A','15453532-17-T/C','15471368-21-G/T','11481364-21-C/G','11474682-27-G/T','15454461-35-C/G','15473867-40-C/A','11499678-10-T/C','72482749-42-G/C','26736425-47-C/T','50825275-51-A/G','68348339-41-C/T','11463640-47-A/C','15470584-7-G/T','11487887-48-T/A','100662769-24-A/G','54740134-65-T/A','26729511-42-C/A','15472688-20-A/C','15471876-27-A/G','15480919-45-T/C','11474749-48-G/A','11467546-14-C/T','15482226-11-C/A','15474767-13-T/G','15473995-30-T/C','11470771-61-G/A','11467005-37-T/A','15467605-66-C/A','54738916-47-A/C','15473548-21-T/C','15477146-8-A/G','15460375-45-G/A','68349596-34-T/A','68349632-61-A/G','15476175-57-T/C','15456911-10-G/A','9481485-5-C/T','15476685-29-T/G'))

#neg 0.10
ten_adult2_neg<- gl.keep.loc(ten_adult, loc.list=c('26743523-7-T/C','11470711-40-A/G','15471299-8-G/C','11470588-23-C/T','11461678-8-A/G','15474872-53-C/T','11460604-39-C/T','15469557-10-A/G','15469320-34-A/C','17478545-29-G/A','17478488-27-G/T','15457388-48-C/G','15479064-48-C/T','15472352-60-A/T','15470094-11-C/A','11460827-17-C/A','11463577-12-A/T','15476949-60-C/T','9498501-47-C/T','15460185-6-C/T','15476861-19-C/T','26745102-51-C/A','11480077-22-G/A','15459695-8-T/C','15459969-60-A/G','15457352-32-T/C','11461291-5-G/C','15453551-57-C/T','15462467-30-A/C','15472012-7-G/A','68349508-11-T/C','15476728-8-G/C','11463708-21-C/T','15457068-40-A/T','26739090-32-C/A','26738349-68-T/G','15468891-21-C/A','26732353-16-A/C','11467825-53-T/G','11473143-42-G/A','11443613-26-T/G','15457681-43-A/G','15472232-43-C/T','15478996-42-G/A','50828880-12-G/T','11471890-18-C/A','15466175-11-G/A','11466799-26-T/G','15453501-16-C/G','15472942-36-A/G','15477872-27-A/C','15464392-14-C/G','15460218-42-A/T','11494737-17-G/T','9483938-41-C/T','11470559-68-C/A','11476451-53-C/T','26737058-65-C/G','26743878-66-C/A','15478941-13-C/T','15464701-22-G/A','11467045-31-C/T','15473452-8-C/G','15475982-32-A/G','15473562-10-G/T','9485195-32-G/A','11477802-42-G/A','9429919-10-C/A','15454850-62-G/T','11470276-40-C/T','11470593-51-C/A','11474978-46-G/C','50826019-43-C/A','11495526-40-G/A','15464738-46-T/G','15461523-14-C/A','15474029-31-G/T','15478876-18-T/C','15474925-40-C/T','11466843-58-G/T','15466253-58-T/C','26723097-63-C/A','15482419-29-G/A','15477031-19-G/T','26730126-57-A/T','11470441-68-G/C','39068390-24-G/T','11501439-30-C/T','17480649-31-T/G','50827961-54-C/T','15472067-20-T/C','15462689-22-T/C','15459634-61-C/A','43487524-16-G/C','11443722-23-C/A','11477753-21-C/A','15475296-11-C/T','50824276-19-G/T','15467473-14-T/G','26732903-5-G/A','9433602-11-C/A','11489117-44-C/A','11460839-68-G/T','11477670-31-A/T','11484583-34-C/A','11490533-61-C/T','9819840-54-C/A','15453901-33-C/A','15479433-5-T/C','26735909-5-C/A','15476111-10-C/T','11500418-12-C/T','15466308-48-C/T','17479105-58-A/G','15470760-25-G/A','15483472-19-C/A','11482462-52-A/G','11472298-35-T/C','15476610-12-C/A','15472078-12-G/C','15449659-25-T/C','26742132-35-C/T','11489162-43-G/A','11470852-14-C/G','15467088-35-A/G','15478938-16-A/T','11463554-14-T/C','15479805-6-G/A','17482461-15-G/A','11460887-21-T/A','11491768-51-G/T','11443009-48-C/G','11470554-21-C/T','15471064-21-T/C','11492286-59-A/T','15460350-52-C/G','50827376-12-C/T','11490185-49-C/G','15472440-20-C/T','15472439-68-C/A','15471785-55-A/G','11467711-17-C/T','15473258-9-A/C','11467965-53-C/T','11467007-66-G/A','15476918-24-C/A','11488190-28-G/T','15467369-54-G/A','11471878-58-A/T','15473453-9-G/T','26739795-60-T/A','11474373-25-G/A','11470346-7-G/A','15462690-37-C/T','11493473-33-C/T','15472633-62-A/T','17483183-14-A/T','17479518-53-G/T','11484571-32-G/A','43565532-13-G/A','15454207-51-A/T','11442895-30-C/T','11496557-6-G/A','26739125-35-C/A','11470857-63-G/A','15481852-32-G/A','15471473-32-C/T','11469770-6-C/T','11469755-8-T/A','15476251-6-T/A','26744887-48-T/C','11485994-35-A/T','11465553-24-C/G','50828873-14-C/A','53148636-38-G/A','15477231-13-G/A','26741380-61-C/G','15468199-11-T/C','15473600-34-T/G','15461471-8-G/C','26733753-42-T/G','11443805-67-G/A','11481363-6-T/A','26745174-62-A/G','11464653-65-G/A','15469304-38-G/T','11469457-24-T/C','26744982-8-A/G','11467674-37-C/T','15466073-67-T/C','26736424-54-T/C','11461901-9-T/C','15462894-13-T/A','26735737-14-C/A','15478887-37-G/A','15468572-37-G/A','15471747-10-C/T','26733017-27-A/G','15470388-61-T/C','11487196-21-A/G','15479118-28-G/C','11491467-8-G/A','11472737-28-T/G','15477511-7-C/T','26749007-9-T/A','15474877-15-C/G','26732274-21-G/A','15463544-5-T/A','11476225-25-C/T','17479620-64-T/C','9482638-46-A/C','11473054-24-T/C','15457665-27-C/A','11488969-11-C/G','15475604-18-A/C','15474193-24-A/G','9494084-18-G/C','26739624-7-C/G','11491742-8-T/A','15473080-6-A/T','15477224-34-C/T','26733662-15-G/C','11499615-11-A/T','17481290-66-T/C','11461725-59-T/C','11460773-32-G/A','26740581-60-C/G','9818882-32-G/C','15475949-14-C/A','26740382-7-G/T','17478679-27-T/C','15475200-29-C/T','15462221-27-T/C','17481402-32-A/T','26729660-12-C/G','9494696-11-A/G','15479624-56-A/T','15461611-13-T/G','15481844-26-T/A'))

#0.05 neg
ten_adult2_neg<- gl.keep.loc(ten_adult, loc.list=c('9498501-47-C/T','26745102-51-C/A','15453551-57-C/T','26738349-68-T/G','11471890-18-C/A','11466799-26-T/G','50826019-43-C/A','26723097-63-C/A','15475296-11-C/T','11490533-61-C/T','17479105-58-A/G','11472298-35-T/C','17483183-14-A/T','11461901-9-T/C','15474193-24-A/G','11499615-11-A/T','15475949-14-C/A','17481402-32-A/T'))

#0.05 pos
ten_adult2_pos<- gl.keep.loc(ten_adult, loc.list=c('15457567-60-T/C','15478027-42-A/T','26735296-20-T/C','15470469-27-T/G','15464910-62-C/T','15476258-41-A/G','15471002-32-A/G','11482155-24-T/C','15477968-19-C/A','9489413-39-A/G','15466554-20-G/A','11472783-25-A/G','15471368-21-G/T','11499678-10-T/C'))
```

```{r}
#relatedness of parent alleles
#join 
sig_adult<- gl.join(ten_adult2_pos, ten_adult2_neg)

#filter only parents
parents<- 
sig_adult<- gl.keep.ind(sig_adult, ind.list = c("A.ten_B.Jewell_6", "A.ten_B.Jewell_3", "A.ten_B.Jewell_8", "A.ten_B.Jewell_7", "A.ten_B.Jewell_2", "A.ten_B.Jewell_5", "A.ten_Parke_2", "A.ten_Parke_5", " A.ten_Parke_4", "A.ten_Switzer_6", "A.ten_Switzer_7", "A.ten_Switzer_2", "A.ten_Switzer_4", "A.ten_Palms_1", "A.ten_Palms_4", "A.ten_Palms_2", "A.ten_Palms_5", "A.ten_Davies_13", "A.ten_Davies_8", "A.ten_Davies_10", "A.ten_Davies_1", "A.ten_Davies_6"))

#grm
grm<- gl.grm(sig_adult)
grm_plot<- gl.plot.heatmap(grm)

dist<- dist(grm)

pcoa<- gl.pcoa(sig_adult)
pca<- gl.pcoa.plot(pcoa, sig_adult)

explained_variance <- c(39.2, 13.9)  # % Variation for PC1 and PC2

pcascore<- as.data.frame(pcoa$scores) |> rownames_to_column(var = "Labels")
pops<- as.data.frame(sig_adult$pop)|> dplyr::rename(pop = `sig_adult$pop`)

pcascore<- cbind(pops, pcascore)

# Plot PCA loadings for PC1 and PC2
pca1<- ggplot(pcascore, aes(x = PC1, y = PC2, colour = pop)) +
  geom_point(size = 3) +  # Add points
  #geom_text(vjust = -0.5, hjust = 0.5) +  # Add labels near points
  labs(x = paste0("PC1 (", explained_variance[1], "%)"), y = paste0("PC2 (", explained_variance[2], "%)"), colour = "Reef") +
  coord_fixed(ratio = explained_variance[2] / explained_variance[1]) +
  theme_Publication() +
  scale_colour_Publication()
pca1

```

```{r}
#write csv of filtered only positive and negative SNPs

#gl.write.csv(ten_adult2_pos, outfile = "ten_adult2_pos.csv", outpath = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr", verbose = NULL)

#gl.write.csv(ten_adult2_neg, outfile = "ten_adult2_neg.csv", outpath = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr", verbose = NULL)

#gl.write.csv(ten_adult2_pos, outfile = "ten_adult2_pos_05.csv", outpath = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr", verbose = NULL)

#gl.write.csv(ten_adult2_neg, outfile = "ten_adult2_neg_05.csv", outpath = "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr", verbose = NULL)
```

```{r}
#get allele freq of all significant SNPs

#A.ten_Palms_5 <- gl.keep.pop(ten_adult2_pos, pop.list= c("A.ten_Palms_5"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_5_alf <- gl.alf(A.ten_Palms_5)
#A.ten_Palms_5_alf<- A.ten_Palms_5_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_5 = alf1) |> dplyr::select(A.ten_Palms_5)

#A.ten_Switzer_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_2_alf <- gl.alf(A.ten_Switzer_2)
#A.ten_Switzer_2_alf<- A.ten_Switzer_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_2 = alf1) |> dplyr::select(A.ten_Switzer_2)

#A.ten_B.Jewell_7 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_7"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_7_alf <- gl.alf(A.ten_B.Jewell_7)
#A.ten_B.Jewell_7_alf<- A.ten_B.Jewell_7_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_7 = alf1) |> dplyr::select(A.ten_B.Jewell_7)

#A.ten_Davies_4 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_4"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_4_alf <- gl.alf(A.ten_Davies_4)
#A.ten_Davies_4_alf<- A.ten_Davies_4_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_4 = alf1) |> dplyr::select(A.ten_Davies_4)

#A.ten_Davies_12 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_12"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_12_alf <- gl.alf(A.ten_Davies_12)
#A.ten_Davies_12_alf<- A.ten_Davies_12_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_12 = alf1) |> dplyr::select(A.ten_Davies_12)

#A.ten_Palms_6 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Palms_6"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_6_alf <- gl.alf(A.ten_Palms_6)
#A.ten_Palms_6_alf<- A.ten_Palms_6_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_6 = alf1) |> dplyr::select(A.ten_Palms_6)

#A.ten_Switzer_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_3_alf <- gl.alf(A.ten_Switzer_3)
#A.ten_Switzer_3_alf<- A.ten_Switzer_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_3 = alf1) |> dplyr::select(A.ten_Switzer_3)

#A.ten_B.Jewell_8 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_8"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_8_alf <- gl.alf(A.ten_B.Jewell_8)
#A.ten_B.Jewell_8_alf<- A.ten_B.Jewell_8_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_8 = alf1) |> dplyr::select(A.ten_B.Jewell_8)

#A.ten_Davies_5 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_5"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_5_alf <- gl.alf(A.ten_Davies_5)
#A.ten_Davies_5_alf<- A.ten_Davies_5_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_5 = alf1) |> dplyr::select(A.ten_Davies_5)

#A.ten_Davies_13 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_13"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_13_alf <- gl.alf(A.ten_Davies_13)
#A.ten_Davies_13_alf<- A.ten_Davies_13_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_13 = alf1) |> dplyr::select(A.ten_Davies_13)

#A.ten_Parke_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Parke_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Parke_1_alf <- gl.alf(A.ten_Parke_1)
#A.ten_Parke_1_alf<- A.ten_Parke_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Parke_1 = alf1) |> dplyr::select(A.ten_Parke_1)

#A.ten_Switzer_4 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_4"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_4_alf <- gl.alf(A.ten_Switzer_4)
#A.ten_Switzer_4_alf<- A.ten_Switzer_4_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_4 = alf1) |> dplyr::select(A.ten_Switzer_4)

#A.ten_B.Jewell_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_1_alf <- gl.alf(A.ten_B.Jewell_1)
#A.ten_B.Jewell_1_alf<- A.ten_B.Jewell_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_1 = alf1) |> dplyr::select(A.ten_B.Jewell_1)

#A.ten_Backnumbers_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Backnumbers_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Backnumbers_1_alf <- gl.alf(A.ten_Backnumbers_1)
#A.ten_Backnumbers_1_alf<- A.ten_Backnumbers_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Backnumbers_1 = alf1) |> dplyr::select(A.ten_Backnumbers_1)

#A.ten_Davies_6 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_6"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_6_alf <- gl.alf(A.ten_Davies_6)
#A.ten_Davies_6_alf<- A.ten_Davies_6_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_6 = alf1) |> dplyr::select(A.ten_Davies_6)

#A.ten_Davies_D1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_D1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_D1_alf <- gl.alf(A.ten_Davies_D1)
#A.ten_Davies_D1_alf<- A.ten_Davies_D1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_D1 = alf1) |> dplyr::select(A.ten_Davies_D1)

#A.ten_Parke_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Parke_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Parke_2_alf <- gl.alf(A.ten_Parke_2)
#A.ten_Parke_2_alf<- A.ten_Parke_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Parke_2 = alf1) |> dplyr::select(A.ten_Parke_2)

#A.ten_Switzer_5 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_5"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_5_alf <- gl.alf(A.ten_Switzer_5)
#A.ten_Switzer_5_alf<- A.ten_Switzer_5_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_5 = alf1) |> dplyr::select(A.ten_Switzer_5)

#A.ten_B.Jewell_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_2_alf <- gl.alf(A.ten_B.Jewell_2)
#A.ten_B.Jewell_2_alf<- A.ten_B.Jewell_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_2 = alf1) |> dplyr::select(A.ten_B.Jewell_2)

#A.ten_Backnumbers_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Backnumbers_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Backnumbers_2_alf <- gl.alf(A.ten_Backnumbers_2)
#A.ten_Backnumbers_2_alf<- A.ten_Backnumbers_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Backnumbers_2 = alf1) |> dplyr::select(A.ten_Backnumbers_2)

#A.ten_Davies_7 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_7"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_7_alf <- gl.alf(A.ten_Davies_7)
#A.ten_Davies_7_alf<- A.ten_Davies_7_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_7 = alf1) |> dplyr::select(A.ten_Davies_7)

#A.ten_Davies_D2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_D2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_D2_alf <- gl.alf(A.ten_Davies_D2)
#A.ten_Davies_D2_alf<- A.ten_Davies_D2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_D2 = alf1) |> dplyr::select(A.ten_Davies_D2)

#A.ten_Parke_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Parke_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Parke_3_alf <- gl.alf(A.ten_Parke_3)
#A.ten_Parke_3_alf<- A.ten_Parke_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_Parke_3 = alf1) |> dplyr::select(A.ten_Parke_3)

#A.ten_Switzer_6 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_6"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_6_alf <- gl.alf(A.ten_Switzer_6)
#A.ten_Switzer_6_alf<- A.ten_Switzer_6_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_6 = alf1) |> dplyr::select(A.ten_Switzer_6)

#A.ten_B.Jewell_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_3_alf <- gl.alf(A.ten_B.Jewell_3)
#A.ten_B.Jewell_3_alf<- A.ten_B.Jewell_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_3 = alf1) |> dplyr::select(A.ten_B.Jewell_3)

#A.ten_Backnumbers_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Backnumbers_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Backnumbers_3_alf <- gl.alf(A.ten_Backnumbers_3)
#A.ten_Backnumbers_3_alf<- A.ten_Backnumbers_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_Backnumbers_3 = alf1) |> dplyr::select(A.ten_Backnumbers_3)

#A.ten_Davies_8 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_8"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_8_alf <- gl.alf(A.ten_Davies_8)
#A.ten_Davies_8_alf<- A.ten_Davies_8_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_8 = alf1) |> dplyr::select(A.ten_Davies_8)

#A.ten_Palms_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Palms_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_1_alf <- gl.alf(A.ten_Palms_1)
#A.ten_Palms_1_alf<- A.ten_Palms_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_1 = alf1) |> dplyr::select(A.ten_Palms_1)

#A.ten_Parke_4 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Parke_4"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Parke_4_alf <- gl.alf(A.ten_Parke_4)
#A.ten_Parke_4_alf<- A.ten_Parke_4_alf |> rownames_to_column() |> dplyr::rename(A.ten_Parke_4 = alf1) |> dplyr::select(A.ten_Parke_4)

#A.ten_Switzer_7 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_7"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_7_alf <- gl.alf(A.ten_Switzer_7)
#A.ten_Switzer_7_alf<- A.ten_Switzer_7_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_7 = alf1) |> dplyr::select(A.ten_Switzer_7)

#A.ten_B.Jewell_4 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_4"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_4_alf <- gl.alf(A.ten_B.Jewell_4)
#A.ten_B.Jewell_4_alf<- A.ten_B.Jewell_4_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_4 = alf1) |> dplyr::select(A.ten_B.Jewell_4)

#A.ten_Davies_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_1_alf <- gl.alf(A.ten_Davies_1)
#A.ten_Davies_1_alf<- A.ten_Davies_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_1 = alf1) |> dplyr::select(A.ten_Davies_1)

#A.ten_Davies_9 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_9"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_9_alf <- gl.alf(A.ten_Davies_9)
#A.ten_Davies_9_alf<- A.ten_Davies_9_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_9 = alf1) |> dplyr::select(A.ten_Davies_9)

#A.ten_Palms_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Palms_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_2_alf <- gl.alf(A.ten_Palms_2)
#A.ten_Palms_2_alf<- A.ten_Palms_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_2 = alf1) |> dplyr::select(A.ten_Palms_2)

#A.ten_Parke_5 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Parke_5"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Parke_5_alf <- gl.alf(A.ten_Parke_5)
#A.ten_Parke_5_alf<- A.ten_Parke_5_alf |> rownames_to_column() |> dplyr::rename(A.ten_Parke_5 = alf1) |> dplyr::select(A.ten_Parke_5)

#A.ten_B.Jewell_5 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_5"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_5_alf <- gl.alf(A.ten_B.Jewell_5)
#A.ten_B.Jewell_5_alf<- A.ten_B.Jewell_5_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_5 = alf1) |> dplyr::select(A.ten_B.Jewell_5)

#A.ten_Davies_2 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_2"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_2_alf <- gl.alf(A.ten_Davies_2)
#A.ten_Davies_2_alf<- A.ten_Davies_2_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_2 = alf1) |> dplyr::select(A.ten_Davies_2)

#A.ten_Davies_10 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_10"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_10_alf <- gl.alf(A.ten_Davies_10)
#A.ten_Davies_10_alf<- A.ten_Davies_10_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_10 = alf1) |> dplyr::select(A.ten_Davies_10)

#A.ten_Palms_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Palms_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_3_alf <- gl.alf(A.ten_Palms_3)
#A.ten_Palms_3_alf<- A.ten_Palms_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_3 = alf1) |> dplyr::select(A.ten_Palms_3)

#A.ten_Switzer_1 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Switzer_1"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Switzer_1_alf <- gl.alf(A.ten_Switzer_1)
#A.ten_Switzer_1_alf<- A.ten_Switzer_1_alf |> rownames_to_column() |> dplyr::rename(A.ten_Switzer_1 = alf1) |> dplyr::select(A.ten_Switzer_1)

#A.ten_B.Jewell_6 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_B.Jewell_6"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_B.Jewell_6_alf <- gl.alf(A.ten_B.Jewell_6)
#A.ten_B.Jewell_6_alf<- A.ten_B.Jewell_6_alf |> rownames_to_column() |> dplyr::rename(A.ten_B.Jewell_6 = alf1) |> dplyr::select(A.ten_B.Jewell_6)

#A.ten_Davies_3 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_3"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_3_alf <- gl.alf(A.ten_Davies_3)
#A.ten_Davies_3_alf<- A.ten_Davies_3_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_3 = alf1) |> dplyr::select(A.ten_Davies_3)

#A.ten_Davies_11 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Davies_11"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Davies_11_alf <- gl.alf(A.ten_Davies_11)
#A.ten_Davies_11_alf<- A.ten_Davies_11_alf |> rownames_to_column() |> dplyr::rename(A.ten_Davies_11 = alf1) |> dplyr::select(A.ten_Davies_11)

#A.ten_Palms_4 <- gl.keep.pop(ten_adult2, pop.list= c("A.ten_Palms_4"), as.pop = "id", mono.rm = FALSE, recalc=TRUE)
#A.ten_Palms_4_alf <- gl.alf(A.ten_Palms_4)
#A.ten_Palms_4_alf<- A.ten_Palms_4_alf |> rownames_to_column() |> dplyr::rename(A.ten_Palms_4 = alf1) |> dplyr::select(A.ten_Palms_4)
```

```{r}
#bind all together

#ad_alf<- cbind(
#A.ten_Davies_1_alf,
#A.ten_Davies_2_alf,
#A.ten_Davies_3_alf,
#A.ten_Davies_4_alf,
#A.ten_Davies_5_alf,
#A.ten_Davies_6_alf,
#A.ten_Davies_7_alf,
#A.ten_Davies_8_alf,
#A.ten_Davies_9_alf,
#A.ten_Davies_10_alf,
#A.ten_Davies_11_alf,
#A.ten_Davies_12_alf,
#A.ten_Davies_13_alf,
#A.ten_Davies_D1_alf,
#A.ten_Davies_D2_alf,
#A.ten_Backnumbers_1_alf,
#A.ten_Backnumbers_2_alf,
#A.ten_Backnumbers_3_alf,
#A.ten_Palms_1_alf,
#A.ten_Palms_2_alf,
#A.ten_Palms_3_alf,
#A.ten_Palms_4_alf,
#A.ten_Palms_5_alf,
#A.ten_Palms_6_alf,
#A.ten_Parke_1_alf,
#A.ten_Parke_2_alf,
#A.ten_Parke_3_alf,
#A.ten_Parke_4_alf,
#A.ten_Parke_5_alf,
#A.ten_B.Jewell_1_alf,
#A.ten_B.Jewell_2_alf,
#A.ten_B.Jewell_3_alf,
#A.ten_B.Jewell_4_alf,
#A.ten_B.Jewell_5_alf,
#A.ten_B.Jewell_6_alf,
#A.ten_B.Jewell_7_alf,
#A.ten_B.Jewell_8_alf,
#A.ten_Switzer_1_alf,
#A.ten_Switzer_2_alf,
#A.ten_Switzer_3_alf,
#A.ten_Switzer_4_alf,
#A.ten_Switzer_5_alf,
#A.ten_Switzer_6_alf,
#A.ten_Switzer_7_alf
#)
```

```{r}
#get mean allele freq for each individual
#ad_alf2<- colMeans(ad_alf) |> as.data.frame() |> dplyr::rename(ad_alf = `colMeans(ad_alf)`) |> rownames_to_column(var = "IID")
```

```{r}
#load adult significant SNP allele frequency and survival deviance data
#ad_alf plot is 0.10 sig
ad_alf<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/alf_corr_plot.csv")
#ad_alf plot 2 is 0.05 sig
ad_alf2<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/alf_corr_plot2.csv")
#Deviance survival (parental contribution)
cov <- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/parental_genotype_corr/aten_adult.PRS2.cov.csv") |> dplyr::select(IID, Deviance)
```

```{r}
#plot 0.10 sig
ad_alf |>
  ggplot() +
  geom_boxplot(aes(x = Reef, y = alf)) +
  theme_Publication() +
  scale_colour_Publication() +
  labs(y = "Allele Frequency")

#plot 0.05 sig
ad_alf2 |>
  ggplot() +
  geom_boxplot(aes(x = Reef, y = alf)) +
  theme_Publication() +
  scale_colour_Publication() +
  labs(y = "Allele Frequency")
```

```{r}
#join the covariate (parental contribution) with allele frequecies of Sig SNPs
#0.10 sig
alf3<- left_join(cov, ad_alf, by = "IID") |> na.omit()
#0.05 sig
alf4<- left_join(cov, ad_alf2, by = "IID") |> na.omit()
```

```{r}
#calcualte corr coefficient for the survival and Allele freq of significant SNPs (minus outlier A.ten_B.Jewell_7)
alf5<- alf3 |> filter(IID != "A.ten_B.Jewell_7")
cor_test_result1 <- cor.test(alf5$Deviance, alf5$alf)
cor_test_result2 <- cor.test(alf4$Deviance, alf4$alf)

# View the correlation coefficient and p-value
ad_cor<- cor_test_result1$estimate  # Correlation coefficient
ad_p<- cor_test_result1$p.value   # P-value

#adults boxplots by reef
alf_point<- alf3 |> filter(IID != "A.ten_B.Jewell_7") |>
  ggplot() +
    geom_point(aes(x=alf, y=Deviance, color=Reef), size = 2) +
    geom_point(aes(x = 0.17876344, y = 2.67747089), color = "#fdb462", size = 2, shape = 21, fill = NA) + #add the outlier in
    #geom_line(stat = "identity") +
    geom_smooth(aes(x=alf, y=Deviance), method=lm, se=FALSE, colour = "black") +
    annotate("text", x=0.17, y=2.4, label= "R = 0.658, p = 0.0008", size = 6) +
    theme_Publication() +
    #scale_color_manual(values=pal[c(1,2,3,4,5)]) +
    scale_colour_Publication() +
    #ylim(0.11,0.13) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  labs(x = "Allele Frequency", y = "Parental ICC")
alf_point

# View the correlation coefficient and p-value
#cor_test_result2$estimate  # Correlation coefficient
#cor_test_result2$p.value   # P-value

#alf_point<- alf4 |> ggplot() +
#    geom_point(aes(x=alf, y=Deviance, color=Reef), size = 2) +
    #geom_line(stat = "identity") +
    #geom_smooth(aes(x=alf, y=Deviance), method=lm, se=FALSE, colour = "black") +
    #theme_Publication() +
    #scale_color_manual(values=pal[c(1,2,3,4,5)]) +
    #scale_colour_Publication() +
    #ylim(0.11,0.13) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  #labs(x = "Allele Frequency")
#alf_point

```
