---
title: "Untitled"
output: html_document
date: "2024-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r}
# Install the latest version of dartR (v2)
#install.packages("dartR")
library(dartR)
library(tidyverse)
library(ggpubr)
library(adegenet)
library(poppr)
library(hierfstat)
library(reshape2)
library(RColorBrewer)
library(scales)
library(LEA)
library(wesanderson)
```

```{r}
theme_Publication <- function(base_size=14, base_family="arial") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

pal <- wes_palette("Zissou1", 5, type = "discrete")
```

```{r}
# Function for summarising standard error
summarySE <- function(data=NULL, measurevar=NULL, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  require(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm==TRUE) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Commenting this rename mean column line, as not assigning measurevar name properly
  # # Rename the "mean" column    
  # datac <- rename(datac, measurevar = mean)
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

set working directory
```{r, echo = FALSE, comment= FALSE, message=FALSE}
#setwd("/Users/alexmacadam/Desktop/A_tenuis/larvae/allele_freq/assoc_files")
```

Fisher model
#####
```{r}
ten_larvae <- readRDS("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ten_larvae_filter.RDS")
#keep T0
ten_larvae_amb<- gl.keep.pop(ten_larvae, pop.list = c('Amb'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)
ten_larvae_hot<- gl.keep.pop(ten_larvae, pop.list = c('Hot'), as.pop = 'Temp', mono.rm = FALSE, recalc=TRUE)

#allele freq central amb
amb_cen<- gl.keep.pop(ten_larvae_amb, pop.list= c("pure_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_cen<- gl.alf(amb_cen) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq hybrid amb
amb_hyb<- gl.keep.pop(ten_larvae_amb, pop.list= c("hybrid_north_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_hyb<- gl.alf(amb_hyb) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq north amb
amb_nor<- gl.keep.pop(ten_larvae_amb, pop.list= c("pure_north"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_amb_nor<- gl.alf(amb_nor) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_amb = alf1, alf2_amb = alf2)

#allele freq central hot
hot_cen<- gl.keep.pop(ten_larvae_hot, pop.list= c("pure_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_cen<- gl.alf(hot_cen) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#allele freq hybrid hot
hot_hyb<- gl.keep.pop(ten_larvae_hot, pop.list= c("hybrid_north_south"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_hyb<- gl.alf(hot_hyb) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#allele freq north hot
hot_nor<- gl.keep.pop(ten_larvae_hot, pop.list= c("pure_north"), as.pop = "pure_hybrid", mono.rm = FALSE, recalc=TRUE)
af_hot_nor<- gl.alf(hot_nor) |> rownames_to_column(var = "SNP") |> dplyr::rename(alf1_hot = alf1, alf2_hot = alf2)

#Join ambient and hot datat for each region cross
af_cen<- left_join(af_amb_cen, af_hot_cen, by = "SNP") |> mutate(Region = "Central")
af_hyb<- left_join(af_amb_hyb, af_hot_hyb, by = "SNP") |> mutate(Region = "Hybrid")
af_nor<- left_join(af_amb_nor, af_hot_nor, by = "SNP") |> mutate(Region = "North")

#combine all data
af<- rbind(af_cen, af_hyb, af_nor)
```

Central
```{r}
# Load necessary library
library(dplyr)

# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_cen2<- af_cen %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_cen2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_cen2)) {
    # Extract counts for each SNP
    count1_allele <- af_cen2$alf1_amb[i]
    count1_non_allele <- af_cen2$alf2_amb[i]
    count2_allele <- af_cen2$alf1_hot[i]
    count2_non_allele <- af_cen2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_cen2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_cen2$p_adj <- p.adjust(af_cen2$p_value, method = "BH")
```

Hybrid
```{r}
# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_hyb2<- af_hyb %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_hyb2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_hyb2)) {
    # Extract counts for each SNP
    count1_allele <- af_hyb2$alf1_amb[i]
    count1_non_allele <- af_hyb2$alf2_amb[i]
    count2_allele <- af_hyb2$alf1_hot[i]
    count2_non_allele <- af_hyb2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_hyb2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_hyb2$p_adj <- p.adjust(af_hyb2$p_value, method = "BH")
```

North
```{r}
# Set total count for each group (e.g., 100 alleles per group)
total_count <- 100

# Convert allele frequencies to counts based on total_count
af_nor2<- af_nor %>%
  mutate(
    alf1_amb = round(alf1_amb * total_count),
    alf2_amb = round(alf2_amb * total_count),
    alf1_hot = round(alf1_hot * total_count),
    alf2_hot = round(alf2_hot * total_count)
  )

# Ensure the data exists and columns are correctly named
if (!all(c("alf1_amb", "alf2_amb", "alf1_hot", "alf2_hot") %in% names(af_nor2))) {
  stop("Column names do not match. Please ensure they are: alf1_amb, alf2_amb, alf1_hot, alf2_hot")
}

# Initialize an empty vector to store p-values
p_vals <- c()

# Loop through each SNP
for (i in 1:nrow(af_nor2)) {
    # Extract counts for each SNP
    count1_allele <- af_nor2$alf1_amb[i]
    count1_non_allele <- af_nor2$alf2_amb[i]
    count2_allele <- af_nor2$alf1_hot[i]
    count2_non_allele <- af_nor2$alf2_hot[i]
    
    # Ensure counts are not NULL
    if (is.null(count1_allele) || is.null(count1_non_allele) || is.null(count2_allele) || is.null(count2_non_allele)) {
        stop("Some allele counts are NULL. Please check your data.")
    }
    
    # Create a 2x2 table for Fisher's test
    contingency_table <- matrix(c(count1_allele, count1_non_allele,
                                  count2_allele, count2_non_allele), nrow = 2)
    
    # Perform Fisher's exact test
    result <- fisher.test(contingency_table)
    
    # Append the p-value to the vector
    p_vals <- c(p_vals, result$p.value)
}

# Store p-values back into the dataframe for reference
af_nor2$p_value <- p_vals

# Apply Benjamini-Hochberg correction
af_nor2$p_adj <- p.adjust(af_nor2$p_value, method = "BH")
```

#extract the significant SNPs
```{r}
sig1<- af_cen2 |> filter(p_adj < 0.05) |> pull(SNP)
sig2<- af_hyb2 |> filter(p_adj < 0.05) |> pull(SNP)
sig<- c(sig1, sig2)
```

No filtering!
```{r}
af2_all<- af |>
  dplyr::select(SNP, Region, alf1_amb, alf1_hot) |> 
  dplyr::rename(amb = alf1_amb, hot = alf1_hot) |>
  pivot_longer(cols=c('amb', 'hot'),
                    names_to='temp',
                    values_to='alf')

# Assuming your original column is called 'alf'
af2_all$alf_new <- ifelse(af2_all$alf > 0.5, 1 - af2_all$alf, af2_all$alf)
```

```{r}
cols<- c("#277da1","#f94144")

af_plot_all<- af2_all |> 
  ggplot() +
  geom_density(aes(alf_new, fill = temp), adjust = 1.2, alpha = 0.5) +
  facet_wrap(~Region, nrow = 1) +
  theme_classic() +
  scale_fill_manual(values=cols) +
  labs(x="Allele Frequency", y="Density")
af_plot_all
```

Filtered for significant SNPs
```{r}
af2_sig<- af |> filter(SNP %in% sig) |>
  dplyr::select(SNP, Region, alf1_amb, alf1_hot) |> 
  dplyr::rename(amb = alf1_amb, hot = alf1_hot) |>
  pivot_longer(cols=c('amb', 'hot'),
                    names_to='temp',
                    values_to='alf')

# Assuming your original column is called 'alf'
af2_sig$alf_new <- ifelse(af2_sig$alf > 0.5, 1 - af2_sig$alf, af2_sig$alf)
```

plot AF
```{r}
af_plot<- af2_sig |> 
  ggplot() +
  geom_density(aes(alf_new, fill = temp), adjust = 1.5, alpha = 0.5) +
  facet_wrap(~Region, nrow = 1) +
  theme_classic() +
  scale_fill_manual(values=cols) +
  labs(x="Allele Frequency", y="Density")
af_plot
```

No difference in the hot, well heat adapted. the hybirds follow this so may have inherited some beneficial heat associated alleles. central has a big difference in the hot treatments. Many allele freqs move to nearer zero. however, some went towards 0.5, which may indicate heterozygote advantage or heterosis. overdominance, where the heterozygous genotype has a higher adaptive value than either homozygous genotype.

KS test for distribution of significant SNPs
```{r}
#central
ks_central<- af2_sig |> filter(Region == "Central") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_central$amb, ks_central$hot)

#hybrid
ks_hybrid<- af2_sig |> filter(Region == "Hybrid") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_hybrid$amb, ks_hybrid$hot)

#north
ks_north<- af2_sig |> filter(Region == "North") |> dplyr::select(SNP, temp, alf) |> pivot_wider(names_from = temp, values_from = alf)
ks.test(ks_north$amb, ks_north$hot)
```

Manhatten plot
```{r}
#create log 10 p values
af_sig<- rbind(af_cen2, af_hyb2, af_nor2)
af_sig<- af_sig |> mutate(`-log10(p)` = -log10(p_value)) |> mutate(`-log10(p)` = coalesce(`-log10(p)`, 0.01))
```

```{r}
positions<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/positions2.csv")

af_sig2<- merge(af_sig, positions, by='SNP', all.x=TRUE) |>
  dplyr::select(CHR, BP, SNP, REF, ALT, alf1_amb, alf2_amb, alf1_hot, alf2_hot, Region, p_value, p_adj, `-log10(p)`) |>
  na.omit()

#write.table(af_sig2, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/af_sig.txt", quote=F, row.names=F, sep = "\t")
```

```{r}
# Chromosome length info
lengths <- read_tsv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/ragtag.scaffold_lengths.txt", col_names = c("chr","length"),show_col_types = FALSE)

offsets <- lengths %>% dplyr::arrange(desc(length)) %>% 
  dplyr::mutate(offset=cumsum(length)-length) %>% 
  dplyr::mutate(scaffold_num = row_number())

axis_chr<- offsets %>% 
  mutate(centre=offset+length/2) %>% 
  mutate(chr_id=ifelse(grepl(chr,pattern="chr"),chr,"Unplaced")) %>% 
  mutate(chr_id=str_remove(chr_id,"_RagTag"),chr_id=str_remove(chr_id,"chr")) %>% 
  group_by(chr_id) %>% 
  dplyr::summarise(centre = mean(centre))
```

### Conversion to pseudo-chromosome coordinates

All outputs were converted to RagTag pseudo-chromosome coordinates.  The process for is illustrated below

In this case the input example has four columns. The first are chromosome and position.  Then there are two more data columns.  You can have as many data columns as you like. 

```{bash}
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data

#python ./translate_coords.py --keep af_sig.txt ragtag.scaffold.agp > af_sig.ragtag.txt
```

```{r}
mh_data_fisher <- read_table("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/af_sig.ragtag.txt",col_names = c("chr","pos","SNP","Ref","Alt","alf1_amb","alf2_amb","alf1_hot","alf2_hot","Region","p_value","p_adj","-log10(p)","CHR2","pos_orig")) %>%
  left_join(offsets) %>% 
  mutate(abs_pos = pos + offset) %>% 
  mutate(chr_type=ifelse(scaffold_num %% 2 == 0,"dark", "light")) %>% 
  mutate(chr_type=ifelse(grepl(chr,pattern="chr"),chr_type,"unplaced"))
```

Summarise number of significant SNPs per chromosome
```{r}
summary <- mh_data_fisher |>
  dplyr::filter(p_adj < 0.05) |>
  group_by(scaffold_num) |>
  dplyr::summarise(count = n())
```

Plot all
```{r}
ccols <- hcl.colors(n=3,palette = "Tropic")

mh_plot_fisher<- mh_data_fisher %>%
  ggplot() + 
  geom_point(aes(x=abs_pos/1e+6,y=`-log10(p)`,color=chr_type),size=1,alpha=.6) + 
  scale_x_continuous(label = axis_chr$chr_id, breaks = axis_chr$centre/1e6) +
  #geom_hline(yintercept = 2.500488, size = 0.6) +
  geom_hline(yintercept = 3.576385, size = 0.6) +
  xlab("Chromosome") +
  ylab("-log10(p)") + 
  theme_pubr() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust=1)) +
  guides(color="none") +
  #scale_color_manual(values = c("light"=ccols[1],"dark"="grey","unplaced"=ccols[3]))
  scale_color_manual(values = c("light"="#e8998d","dark"="#6c9a8b","unplaced"="darkgrey"))
mh_plot_fisher
```

plot by region
```{r}
mh_plot_region<- mh_data_fisher %>%
  ggplot() + 
  geom_point(aes(x=abs_pos/1e+6,y=`-log10(p)`,color=chr_type),size=1,alpha=.6) + 
  scale_x_continuous(label = axis_chr$chr_id, breaks = axis_chr$centre/1e6) +
  #geom_hline(yintercept = 2.5) +
  xlab("Chromosome") +
  ylab("-log10(p)") + 
  theme_pubr() +
  guides(color="none") +
  facet_wrap(~Region) +
  scale_color_manual(values = c("light"=ccols[1],"dark"="grey","unplaced"=ccols[3]))
mh_plot_region
```

#######
Find genes that SNPs are in
```{r}
#select only the SNP column
sig_SNPs<- mh_data_fisher |> dplyr::filter(p_adj < 0.05) |> dplyr::select(SNP) |> distinct()

#remove header
names(sig_SNPs)= NULL

#write out list
#write.table(sig_SNPs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting/sig_SNPs.txt", row.names = FALSE, quote = F)

#edit the SNPs to the vcf format
```

```{bash}
#create vcf with only significant SNPs
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting
#vcftools --vcf aten_larvae_T0.vcf --snps sig_SNPs.txt --recode --out filtered.vcf
```

```{bash}
#extract the info from the filtered vcf file
#change directory
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting
#save vcf as object
#VCF=/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting/filtered.vcf
#extract first 3 columns
#grep -v "^#" $VCF | cut -f1-3 | awk '{print $0"\t"NR}' > larv_SNPs.txt
```

```{bash}
#convert filtered vcf file to bed files
#change directory
#cd /Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting
#with bedops, convert vcf to bed file
#vcf2bed --snvs < filtered.vcf > filtered.bed
#save bed file as object
#BED=/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/manhattan_plotting/filtered.bed
#extract first 4 columns
#grep -v "^#" $BED | cut -f1-4 | awk '{print $0"\t"NR}' > sig_SNPs.bed
#with bedops, convert gtf file to bed file of genes
#gtf2bed < aten.gtf > genes.bed
```

Flanking
```{bash}
#extract gene names that SNPs are within 10kb of
#Use BEDTools flank command to extend the SNP regions by 10 kb on both sides
#bedtools flank -i sig_SNPs.bed -g genes.bed -b 10000 > snps_10k.bed
#once you have both the SNP BED file and the gene BED file, you can use bedtools intersect to find overlaps
#bedtools intersect -a snps_10k.bed -b genes.bed > snps_flanking.bed
#you can extract the gene names as well into the bed file
#bedtools intersect -a snps_10k.bed -b genes.bed -wo > snps_genes_flanking.bed
#extract SNPs and gene names
#cut -f 4,8,9 snps_genes_flanking.bed > snps_flanking_gene_names.txt
#extract only gene names
#cut -f 3 snps_flanking_gene_names.txt > flanking_gene_names.txt
#keep only unique gene names
#sort flanking_gene_names.txt | uniq > flanking_gene_names_unique.txt
#extract protein fastas of each gene
#seqkit grep -f flanking_gene_names_unique.txt aten_0.11.maker_post_001.proteins.fasta -o sig_flanking_gene_seqs.fasta
```

Use flanking genes and see which
```{r}
#load tsv
annotation<- read.csv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/annotation_table.csv")
#load significant SNP flanking genes list
sig<- read_tsv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/af_flanking_gene_names_unique.txt", col_names = F)
#filter only significant genes
sig_ann <- annotation %>% filter(aten_id2 %in% sig$X1)
#which ones aren't in there?
missing_ann <- sig %>% filter(!X1 %in% sig_ann$aten_id)
```

KOG
```{r}
library(KOGMWU)
library(tidyverse)
#load gene2kog data
gene2kog_aten<- read_tsv("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/Data/Atenuis_gene2kogClass.tab", col_names = F) |> na.omit() 
gene2kog_aten<- gene2kog_aten |> mutate(
  V1 = factor(X1),
  V2 = factor(X2)
) |>
  dplyr::select(V1, V2) |> as.data.frame()
```

```{r}
#significant SNPs data
sig_SNPs1<- mh_data_fisher |> mutate(r = alf1_amb - alf1_hot) |> dplyr::select(SNP, r, p_adj)

snps_flanking_gene_names_m1<- read.table("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/af_snps_flanking_gene_names(DART).txt", header = F) |>
  dplyr::rename(SNP = V1, gene = V3)

sig_SNPs3<- sig_SNPs1 |> left_join(snps_flanking_gene_names_m1, by = "SNP") |>
  dplyr::select(gene, r) |>
  na.omit() |>
  distinct(gene, .keep_all=TRUE) |>
  as.data.frame() |>
  mutate(gene = as.factor(gene))
```

```{r}
# Analyzing D1 coral response to heat stress:
KOG=kog.mwu(sig_SNPs3,gene2kog_aten)
KOG
#write.csv(KOG, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/Final analysis/af_KOG_delta_ranks.csv")
```

Plot KOG terms
```{r}
KOG_plot2<- ggplot(data = KOG, aes(x = delta.rank, y = reorder(term, delta.rank), color = padj, size = nseqs)) +
  geom_point() +
  scale_color_gradient(low = "red", high = "blue") +
  labs(x = "Delta Rank", y = "KOG Term", color = "Adjusted P-value", size = "N Seqs") +
  theme_minimal()
KOG_plot2

KOG_plot1<- ggplot(data = KOG, aes(x = reorder(term, nseqs), y = nseqs, fill = delta.rank)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_gradient(low = "red", high = "blue") +
  facet_wrap(~ cut(padj, breaks = c(0, 0.05, 0.1, 1), labels = c("p <= 0.05", "0.05 < p <= 0.1", "p > 0.1"))) +
  labs(x = "KOG Term", y = "Number of Sequences", fill = "Delta rank") +
  theme_minimal()
KOG_plot1

KOG_plot3<- ggplot(data = KOG, aes(x = delta.rank, y = reorder(term, delta.rank), color = nseqs)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "red", high = "blue") +
  labs(x = "Delta Rank", y = "KOG Term", color = "N seqs") +
  theme_minimal()
KOG_plot3

KOG_plot<- KOG_plot1 + KOG_plot2
KOG_plot
```

GO enrichment analysis
```{r}
#read in sig genes
#find gene names for sig SNPs
sig_SNPs_GO<- sig_SNPs1 |> left_join(snps_flanking_gene_names_m1, by = "SNP") |>
  dplyr::select(gene, r, p_adj) |>
  distinct(gene, .keep_all=TRUE) |>
  as.data.frame() |>
  mutate(gene = as.factor(gene)) |>
  na.omit()

sig_SNPs_GO$p_adj[is.na(sig_SNPs_GO$p_adj)] <- 1

#extract gene names
genes<- gene2kog_aten |> dplyr::select(V1) |> dplyr::rename(gene=V1)
#join all gene names to significant genes
sig_genes2<- genes |> left_join(sig_SNPs_GO, by = "gene")
#add .m1 to gene names to correspond with other datasets
sig_genes2$gene <- paste0(sig_genes2$gene, ".m1")
#turn non-sig genes NAs to 0s
sig_genes2$r[is.na(sig_genes2$r)] <- 0
#select relevant columns
sig_genes3<- sig_genes2 |> dplyr::select(gene, r)
#write dataframe
#write.csv(sig_genes3, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO/sig_genes.csv")

#genes of only the significant SNPs
sig_genes4<- sig_genes2 |> dplyr::filter(p_adj < 0.05) |> dplyr::select(gene, r)
#write dataframe
#write.csv(sig_genes4, "/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO/sig_genes_only.csv")
```

```{r}
#All genes
# First, press command-D on mac or ctrl-shift-H in Rstudio and navigate to the directory containing scripts and input files. Then edit, mark and execute the following bits of code, one after another.
setwd("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO")
getwd()

#NOTE REMEMBER!!!! add '.m1' to the end of the gene name to make it run

# Edit these to match your data file names: 
input="sig_genes.csv" # two columns of comma-separated values: gene id, continuous measure of significance. To perform standard GO enrichment analysis based on Fisher's exact test, use binary measure (0 or 1, i.e., either sgnificant or not).
#input="all_genes.csv" # all genes version
goAnnotations= "atenuis_gene2go.tab" # two-column, tab-delimited, one line per gene, multiple GO terms separated by semicolon. If you have multiple lines per gene, use nrify_GOtable.pl prior to running this script.
goDatabase="go.obo" # download from http://www.geneontology.org/GO.downloads.ontology.shtml
goDivision="MF" # either MF, or BP, or CC
source("/Users/alexmacadam/Dropbox/PhD/Chapter4_RNA/Data/Enrichment_analysis/Larvae/gomwu.functions.R")

# ------------- Calculating stats
# It might take a few minutes for MF and BP. Do not rerun it if you just want to replot the data with different cutoffs, go straight to gomwuPlot. If you change any of the numeric values below, delete the files that were generated in previos runs first.
 
gomwuStats(input, goDatabase, goAnnotations, goDivision,
	perlPath="perl", # replace with full path to perl executable if it is not in your system's PATH already
	largest=0.2,  # a GO category will not be considered if it contains more than this fraction of the total number of genes
	smallest=2,   # a GO category should contain at least this many genes to be considered
	clusterCutHeight=0.25, # threshold for merging similar (gene-sharing) terms. See README for details.
#	Alternative="g" # by default the MWU test is two-tailed; specify "g" or "l" of you want to test for "greater" or "less" instead. 
#	Module=TRUE,Alternative="g" # un-remark this if you are analyzing a SIGNED WGCNA module (values: 0 for not in module genes, kME for in-module genes). In the call to gomwuPlot below, specify absValue=0.001 (count number of "good genes" that fall into the module)
#	Module=TRUE # un-remark this if you are analyzing an UNSIGNED WGCNA module 
)
# do not continue if the printout shows that no GO terms pass 10% FDR.

# ----------- Plotting results
quartz()
results=gomwuPlot(input,goAnnotations,goDivision,
 	absValue=-log(0.05,10),  # genes with the measure value exceeding this will be counted as "good genes". This setting is for signed log-pvalues. Specify absValue=0.001 if you are doing Fisher's exact test for standard GO enrichment or analyzing a WGCNA module (all non-zero genes = "good genes").
 #	absValue=1, # un-remark this if you are using log2-fold changes
 	level1=0.1, # FDR threshold for plotting. Specify level1=1 to plot all GO categories containing genes exceeding the absValue.
 	level2=0.05, # FDR cutoff to print in regular (not italic) font.
 	level3=0.01, # FDR cutoff to print in large bold font.
 	txtsize=1.2,    # decrease to fit more on one page, or increase (after rescaling the plot so the tree fits the text) for better "word cloud" effect
 	treeHeight=0.5, # height of the hierarchical clustering tree
 #	colors=c("dodgerblue2","firebrick1","skyblue2","lightcoral") # these are default colors, un-remar and change if needed
 )
 # manually rescale the plot so the tree matches the text 
 # if there are too many categories displayed, try make it more stringent with level1=0.05,level2=0.01,level3=0.001.  
 
 # text representation of results, with actual adjusted p-values
 res<-results[[1]]
write.csv(res,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO/GO_Aten.csv")
# ------- extracting representative GOs

# this module chooses GO terms that best represent *independent* groups of significant GO terms

pcut=1e-2 # adjusted pvalue cutoff for representative GO
hcut=0.9 # height at which cut the GO terms tree to get "independent groups". 

# plotting the GO tree with the cut level (un-remark the next two lines to plot)
quartz()
 plot(results[[2]],cex=0.6)
 abline(h=hcut,col="red")

# cutting
ct=cutree(results[[2]],h=hcut)
annots=c();ci=1
for (ci in unique(ct)) {
  message(ci)
	rn=names(ct)[ct==ci]
	obs=grep("obsolete",rn)
	if(length(obs)>0) { rn=rn[-obs] }
	if (length(rn)==0) {next}
	rr=results[[1]][rn,]
	bestrr=rr[which(rr$pval==min(rr$pval)),]
	best=1
	if(nrow(bestrr)>1) {
		nns=sub(" .+","",row.names(bestrr))
		fr=c()
		for (i in 1:length(nns)) { fr=c(fr,eval(parse(text=nns[i]))) }
		best=which(fr==max(fr))
	}
	if (bestrr$pval[best]<=pcut) { annots=c(annots,sub("\\d+\\/\\d+ ","",row.names(bestrr)[best]))}
}

mwus=read.table(paste("MWU",goDivision,input,sep="_"),header=T)
bestGOs=mwus[mwus$name %in% annots,]
bestGOs

#write.csv(mwus,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_mwus.csv")
#write.csv(bestGOs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_bestGOs.csv")
```

```{r}
#Sig genes
# First, press command-D on mac or ctrl-shift-H in Rstudio and navigate to the directory containing scripts and input files. Then edit, mark and execute the following bits of code, one after another.
setwd("/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO")
getwd()

#NOTE REMEMBER!!!! add '.m1' to the end of the gene name to make it run

# Edit these to match your data file names: 
input="sig_genes_only.csv" # two columns of comma-separated values: gene id, continuous measure of significance. To perform standard GO enrichment analysis based on Fisher's exact test, use binary measure (0 or 1, i.e., either sgnificant or not).
#input="all_genes.csv" # all genes version
goAnnotations= "atenuis_gene2go.tab" # two-column, tab-delimited, one line per gene, multiple GO terms separated by semicolon. If you have multiple lines per gene, use nrify_GOtable.pl prior to running this script.
goDatabase="go.obo" # download from http://www.geneontology.org/GO.downloads.ontology.shtml
goDivision="MF" # either MF, or BP, or CC
source("/Users/alexmacadam/Dropbox/PhD/Chapter4_RNA/Data/Enrichment_analysis/Larvae/gomwu.functions.R")

# ------------- Calculating stats
# It might take a few minutes for MF and BP. Do not rerun it if you just want to replot the data with different cutoffs, go straight to gomwuPlot. If you change any of the numeric values below, delete the files that were generated in previos runs first.
 
gomwuStats(input, goDatabase, goAnnotations, goDivision,
	perlPath="perl", # replace with full path to perl executable if it is not in your system's PATH already
	largest=0.2,  # a GO category will not be considered if it contains more than this fraction of the total number of genes
	smallest=2,   # a GO category should contain at least this many genes to be considered
	clusterCutHeight=0.25, # threshold for merging similar (gene-sharing) terms. See README for details.
#	Alternative="g" # by default the MWU test is two-tailed; specify "g" or "l" of you want to test for "greater" or "less" instead. 
#	Module=TRUE,Alternative="g" # un-remark this if you are analyzing a SIGNED WGCNA module (values: 0 for not in module genes, kME for in-module genes). In the call to gomwuPlot below, specify absValue=0.001 (count number of "good genes" that fall into the module)
#	Module=TRUE # un-remark this if you are analyzing an UNSIGNED WGCNA module 
)
# do not continue if the printout shows that no GO terms pass 10% FDR.

# ----------- Plotting results
quartz()
results=gomwuPlot(input,goAnnotations,goDivision,
 	absValue=-log(0.05,10),  # genes with the measure value exceeding this will be counted as "good genes". This setting is for signed log-pvalues. Specify absValue=0.001 if you are doing Fisher's exact test for standard GO enrichment or analyzing a WGCNA module (all non-zero genes = "good genes").
 #	absValue=1, # un-remark this if you are using log2-fold changes
 	level1=0.1, # FDR threshold for plotting. Specify level1=1 to plot all GO categories containing genes exceeding the absValue.
 	level2=0.05, # FDR cutoff to print in regular (not italic) font.
 	level3=0.01, # FDR cutoff to print in large bold font.
 	txtsize=1.2,    # decrease to fit more on one page, or increase (after rescaling the plot so the tree fits the text) for better "word cloud" effect
 	treeHeight=0.5, # height of the hierarchical clustering tree
 #	colors=c("dodgerblue2","firebrick1","skyblue2","lightcoral") # these are default colors, un-remar and change if needed
 )
 # manually rescale the plot so the tree matches the text 
 # if there are too many categories displayed, try make it more stringent with level1=0.05,level2=0.01,level3=0.001.  
 
 # text representation of results, with actual adjusted p-values
 res<-results[[1]]
write.csv(res,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/allele_freq/R/GO/GO_Aten.csv")
# ------- extracting representative GOs

# this module chooses GO terms that best represent *independent* groups of significant GO terms

pcut=1e-2 # adjusted pvalue cutoff for representative GO
hcut=0.9 # height at which cut the GO terms tree to get "independent groups". 

# plotting the GO tree with the cut level (un-remark the next two lines to plot)
quartz()
 plot(results[[2]],cex=0.6)
 abline(h=hcut,col="red")

# cutting
ct=cutree(results[[2]],h=hcut)
annots=c();ci=1
for (ci in unique(ct)) {
  message(ci)
	rn=names(ct)[ct==ci]
	obs=grep("obsolete",rn)
	if(length(obs)>0) { rn=rn[-obs] }
	if (length(rn)==0) {next}
	rr=results[[1]][rn,]
	bestrr=rr[which(rr$pval==min(rr$pval)),]
	best=1
	if(nrow(bestrr)>1) {
		nns=sub(" .+","",row.names(bestrr))
		fr=c()
		for (i in 1:length(nns)) { fr=c(fr,eval(parse(text=nns[i]))) }
		best=which(fr==max(fr))
	}
	if (bestrr$pval[best]<=pcut) { annots=c(annots,sub("\\d+\\/\\d+ ","",row.names(bestrr)[best]))}
}

mwus=read.table(paste("MWU",goDivision,input,sep="_"),header=T)
bestGOs=mwus[mwus$name %in% annots,]
bestGOs

#write.csv(mwus,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_mwus.csv")
#write.csv(bestGOs,"/Users/alexmacadam/Dropbox/PhD/Chapter3_Genomics/A_tenuis/04_larvae/alfreq_surv_correlation/corr_version/GO_Aten_bestGOs.csv")
```

Plot
```{r}
library(patchwork)
fisher_plots<- af_plot + mh_plot + KOG_plot3 + plot_layout(nrow = 3, heights = c(1,1,1.5)) + plot_annotation(tag_levels = 'A')
fisher_plots
```

